### 2014-03-19 hera-concept2013.tex
initial commit

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for the following class files: copernicus.cls, copernicus2.cls, copernicus_discussions.cls
%% The class files, the Copernicus LaTeX Manual with detailed explanations regarding the comments
%% and some style files are bundled in the Copernicus Latex Package which can be downloaded from the different journal webpages.
%% For further assistance please contact the Publication Production Office (production@copernicus.org).
%% http://publications.copernicus.org

%% Differing commands regarding the specific class files are highlighted.

%% copernicus.cls
\documentclass[ars]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus}

%% copernicus2.cls
%\documentclass[journal abbreviation]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus2}

%% copernicus_discussions.cls
%\documentclass[ars, hvmath, online]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus_discussions}

\def\kperp{k_{\bot}}
\def\kpar{k_{\|}}
\def\nwnh{{\sl NWNH}}

\newcommand{\simgt}{\stackrel{>}{_{\sim}}}
\def\kperp{k_{\bot}}
\def\kpar{k_{\|}}
\def\k{{\bf k}}
\def\sky{{\theta}}
\def\HI{{H{\small I }}}
\def\HII{{H{\small II }}}
\def\xHI{{x_{\rm\HI}}}
\begin{document}

\linenumbers

\title{The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA)}

\author[1]{David R. DeBoer}
\author[2]{James Aguirre}
\author[3]{Judd Bowman}
\author[4]{Richard Bradley}
\author[4]{Chris Carilli}
\author[5]{Josh Dillon}
\author[6]{Steve Furlanetto}
\author[5]{Jacqueline Hewitt}
\author[3]{Daniel Jacobs}
\author[1]{Adrian Liu}
\author[7]{Miguel Morales}
\author[1]{Aaron Parsons}
\author[7]{Jonathan Pober}
\author[5]{Max Tegmark}
\author[1]{Dan Werthimer}

%% each address must have a unique identifier in the option field
\affil[1]{University of California, Berkeley, CA USA}
\affil[2]{University of Pennsylvania, Philadelphia, PA USA}
\affil[3]{Arizona State University, Tempe, AZ USA}
\affil[4]{National Radio Astronomy Observatory, Charlottesville, VA USA}
\affil[5]{Massachusetts Institute of Technology, Cambridge, MA USA}
\affil[6]{University of California, Los Angeles, CA USA}
\affil[7]{University of Washington, Seattle, WA USA}

%% The [] brackets identify the author to the corresponding affiliation, 1, 2, 3, etc. should be inserted.

\runningtitle{HERA}

\runningauthor{DeBoer {\em et al}}

\correspondence{David R. DeBoer\\ (ddeboer@berkeley.edu)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by the Publication Production Office during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged
experiment that uses the unique properties of the 21-cm line from neutral
hydrogen to probe the Epoch of Reionization (EoR) and the preceding Dark
Ages. During these epochs, roughly 0.3-1 Gyr after the Big Bang, the first
stars and black holes heat and reionize the Universe following cosmic
recombination. Direct observation of the large scale structure of
reionization and its evolution with time will have
a profound impact on our understanding of the birth of the first galaxies
and black holes, their influence on the intergalactic medium (IGM), and
cosmology.  HERA is a compact hexagonal-packed array of 14-meter paraboloids 
operating from 50 - 225 MHz.
\end{abstract}

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged project
that uses the unique properties of the 21-cm line of neutral hydrogen to probe the
Epoch of Reionization (EoR) and the preceding Dark Ages. During these epochs, roughly
0.3-1 Gyr after the Big Bang, the first stars and black holes heated and reionized
the Universe, which at that time existed as a nearly uniform sea of warm neutral
hydrogen.   This period of cosmic reionization represents one of the few global
phase changes in the evolution of the Universe.
Direct observation of the large scale structure of reionization and its
evolution with time will have a profound impact on our understanding of the birth of
the first galaxies and black holes, their influence on the intergalactic medium
(IGM), and cosmology.

Figure \ref{fig:theUniverse} is a cartoon showing cosmic evolution from the Big Bang on the left through
to HERA today.

Detecting, characterizing and ultimately imaging this epoch is a key goal in
furthering our understanding of the Universe and was the top priority in the Radio,
Millimeter, and Sub-millimeter category of recommended new facilities in the most
recent decadal survey of astronomical priorities (\ref{NWNH}). Current projects
(PAPER, MWA, LOFAR, GMRT YYYYYGIVE REFSYYYYY) are striving to make the first detection of the statistical
power spectrum of the signal, but current best limits still fall above even
optimistic predictions of its intrinsic strength. While these projects are still
taking data, it is recognized that an optimized array based on our new understanding
of the signal characteristics is needed to make a strong detection and begin to
characterize this signal over multiple scales and redshifts.

HERA overall is a phased set of experiments that build on its predecessors and preceding stages.
The first phase of the HERA roadmap entailed the operation of the
PAPER and MWA telescopes to explore techniques and designs required to
detect the primordial HI signal in the presence of radio continuum
foreground emission some four orders of magnitude brighter. Studies
with PAPER and the MWA have led to a new understanding of the
interplay of foreground and instrumental systematics in the context of
a three-dimensional cosmological intensity-mapping experiment.  YYYYYGIVE REFSYYYYY 
We are
now able to remove foregrounds to the limits of our sensitivity with
these instruments, culminating in the first physically meaningful
upper limits on the power spectrum of 21~cm emission from
reionization (see Fig \ref{fig:eor_pspec}).

Building on this understanding, the next stage of HERA entails a new
14m diameter antenna element that is optimized both for sensitivity
and for minimizing foreground systematics.  Arranging these elements
in a compact hexagonal grid yields an array that facilitates
calibration, leverages proven foreground removal techniques, and is
scalable to large collecting areas. The array will be located near the current PAPER
experiment
in the radio quiet environment of the SKA site in Karoo, South Africa,
and will have a sensitivity close to two orders of magnitude better than
PAPER and the MWA.  HERA's sensitivity coupled with broader frequency coverage, 
will enable HERA to paint an
uninterrupted picture through reionization, back to the end of the
Dark Ages.  As an array, it can be used in smaller arrays along the way to 
achieve timely science.  Two benchmarks are planned:

HERA~127, deployed in 2016, will measure the rise and fall of the
21~cm reionization power spectrum, constraining the timing and
duration of reionization.

HERA~331, deployed in 2017, will measure fluctuations in the 21~cm
signal over a variety of spatial scales to determine the nature and
distribution of the first galaxies that dominate cosmic
reionization. HERA~331 will also extend precision power-spectrum
observations back to the end of the 'Dark Ages' (z $\sim$ 20), when
the first stars and black holes warm the primordial IGM.

This paper will present a summary of the current understanding
of the signal characteristics and measurements and describe this planned
HERA telescope to be built to detect and characterize the EoR power
spectrum.

\section{Background}
\label{sec:background}
\subsection{Hubble's Law}
\label{sec:hubble}
Modern cosmology has provided a spectacular understanding of the cosmic history of
our Universe starting from the instantaneous moment after the explosion of our
spacetime that we call ``the Big Bang'', which occurred about 13.8 Gyr ago, to the
advent of the large structures that seeded our cosmic current structure. After the
Big Bang, matter and space went streaming outward and is manifest to us by the simple
expression called Hubble's Law:
\begin{equation}
v = H_od
\end{equation}
where $v$ is the velocity of the object away from us, $d$ is the distance to the
object $H_o$ is the so-called ``Hubble Constant'', which has a value of 68 Mpc/km/sec
\ref{planck}, in astronomer's units, where Mpc denotes megaparsecs, where a parsec is
$3.086 \times 10^13$ km. Using measured Doppler shifts one may measure $v$, so if one
knows $H_o$ one may directly compute the distance to the object. Velocity is
typically given as a ``redshift'', denoted as $z$, related to one another by the
non-relativistic Doppler shift equation $z \sim v/c$, where $c$ is the speed of light
in vacuo. And then by Hubble's Law, that redshift conveys a distance.

Note that Hubble's ``Constant'' is a function of time and using $H_o$ denotes today's
value. Determining the evolution of Hubble's constant is essentially the role of
cosmological studies and using Hubble's law for very distant sources requires an
integration with cosmological evolution encoded within it. The values here are
derived from \ref{ned_evo}.

\subsection{Power Spectrum}
\label{sec:pspec}
One tool that is typically employed in studying large scale parameters of the
Universe, is to collapse the richness of its spatial variation to its spatial Fourier
transform and compute a power spectrum of this. It is the parameters of this power
spectrum that may be predicted by cosmology, rather than the essentially random
spatial variation about it. It is analogous to the fact that one may get a few broad
parameters of an FM signal by Fourier transforming its time series, regardless of whether it is
Beethoven or M\"{o}tley Cr\"{u}e, which have very different values as a function of time.

Cosmology typically encodes this data as power in spherical harmonics.  Epoch of
Reionization studies typically use wavevectors, $\k$.  This 3-D vector may be decomposed
into plane-of-sky values (${\bf \kperp}$) and a line-of-sight term (${\kpar}$).  The overall
wavevector may then be written as $\k = {\bf \kperp} + \kpar{\bf z}$.  As a wavevector,
the units of $\k$ are 1/Mpc (inverse length).  Note that although modern cosmology believes
it has pegged the value of Hubble's constant, to hedge their bets, they will often factor out
$h=H_o/100$ as an overall scale adjustment, such that the units are $h$/Mpc.  The units
of the power spectrum magnitudes are given in volume-adjusted power-squared (mK$^2$), denoted
$\Delta ^2$.

\subsection{Foregrounds and Overall Signal Levels}
\label{sec:foregrounds}
At the frequencies of interest (50-250 MHz), the sky as seen by a radio telescope is dominated 
by the synchrotron emission from free electrons spiralling around the magnetic fields that thread
through our Galaxy.  The magnitude of this signal varies as a power law of the frequency, as well
as position on the sky (hottest in the plane of the galaxy, cooler at the poles).  The variation spans
from about 10,000K to about 100K in equivalent temperature.  There are a few localized sources
in the sky that are even hotter to the eyes of a radio telescope.

As will be shown shortly, the signals of interest have equivalent temperatures of $<$ 10 mK, 7 to 9 
orders of magnitude below the foreground signal strengths.  The challenge is to peer through these
strong foreground signals to see the weak signal of interest below.

As stated above, the strong foreground signals may be accurately described as smooth power laws in 
frequency - a factor that may be exploited to extract the signals of interest, which have a structure
that varies fairly rapidly with frequency.

It is also interesting to note that another ``foreground'' from our perspective is the Cosmic Microwave Background.

\section{Scientific Motivation}
\label{sec:science}

The {\it cosmic dawn}, the period beginning with the birth of the first stars and
culminating with the full reionization of the IGM some 500 Myrs later, represents one
of the last unexplored phases in the history of structure formation. During this
period, a wealth of astrophysical and cosmological phenomena are at work. The
characteristics of the IGM depend on the cosmic density field, the formation sites of
the first luminous sources (e.g., their typical masses and clustering), their
constituents (e.g., exotic Population III stars, more normal stars, stellar remnants,
or supermassive black holes) their ultraviolet luminosities (which affect the IGM's
ionization state), the efficiency and abundance of X-ray sources (which affect the
IGM temperature), and even more exotic effects like the relative velocity of baryons
and dark matter.

A season of observing with HERA-127 will yield high-significance constraints on the
21 cm power spectrum across a wide range of k modes and redshifts
\citep{pober_et_al2014}. In Figure \ref{fig:eor_pspec} we show the $z=9$ power
spectrum predicted by the publicly available 21cmFAST software
\citep{mesinger_et_al2011}, along with $2\sigma$ HERA sensitivities. Using the
conservative delay-spectrum (``foreground avoidance") approach pioneered by PAPER
(Parsons et al. 2014), we find that HERA-127 can achieve a $> 10\sigma$ detection of
fiducial power spectra over a broad range of redshifts. The subsequent observing
season with HERA-331 can increase this detection significance to over $25\sigma$
using the same methods. With detailed foreground modeling, the more sophisticated
power spectrum estimator developed for the MWA could increase the size of the ``EoR
window", the region of Fourier space with minimal foreground contamination. This
would allow for an overall detection significance of up to $90\sigma$, along with
access to lower $k$ modes and therefore qualitatively different physics. Such a high
sensitivity measurement would also allow one to go beyond constraining parameters,
testing rather than assuming the underlying theoretical framework and starting to
image the large neutral bubbles during reionization.

%
%\begin{figure}[t]\centering
%\includegraphics[height=2.5in]{HERA_331_z8_SNR_annotated.jpg}
%\caption{\small
%With sensitivity highly concentrated at the largest scales, 
%HERA-331 will have the sensitivity to directly image HI during reionization.  Shown here is a simulation of EoR emission \citep{mcquinn_et_al2007} as imaged by HERA in 100 hours of observation.  
%Contours enclose regions with signal to noise above 10.  
%With a nearly completely sampled aperture over 300~m across, HERA will have the collecting area of Arecibo but with $500\times$ the survey speed. Each night, HERA drift scans 2600 square degrees.
%}  \label{fig:imaging}
%\end{figure}

The ability of HERA to image enables an exciting range of cross-correlation science.
HERA HI images reveal the large-scale reionization environment for pointed ALMA and
JWST observations, and other deep near-IR surveys \citep{lidz_et_al2009}. Knowing
whether an observed galaxy is in a region that was previously reionized (center of
large HII bubble), recently reionized (edge of HII bubble), or is forming from
pristine neutral gas provides important contextual information on early galaxy
formation.

The HI images can be cross-correlated with other diffuse tracers of large scale
structure. A number of studies have proposed cross-correlation with large scale
intensity mapping of molecular CO \cite{lidz_et_al2011} and atomic CII
\citep{gong_et_al2011} lines. These studies trace the large scale galaxy distribution
-- the sources of reionization. Prototypes of these experiments may be operating on
the HERA timescale. Such probes have the advantage of having different systematics
compared to HERA, potentially allowing clean measurements of the underlying signal.

VERY SHORT DISCUSSION OF WHAT LIMITS THE EOR BETWEEN Z~12-~6

\section{Specifications}
\label{sec:spec}
The epoch of interest, the spatial scales of interest, the required sensitivity and
the frequency smoothness provide the scientific specifications needed for the array.
Details of the technique, cost, and observing strategies also provide specifications.
This section will discuss these high-level specifications.

\subsection{Frequency Range}
The rest frequency ($\nu_o$) of the hyperfine transition of neutral hydrogen is at
1420.4 MHz. This frequency gets red-shifted to lower frequencies in deep astronomical
observations due to the expansion of the Universe, as discussed above. This is
characterized by the red-shift parameter $z$ as
\begin{equation}
\centering
\label{eq:redshift}
\nu_s(z) = \nu_o/(1+z)
\end{equation}
The age of the Universe at a particular redshift is given by the cosmological model
and a key goal of cosmology is to determine those parameters, however existing models
allow us to determine these values to our desired accuracy. Figure
\ref{fig:theUniverse} plots the red-shifted frequency as a function of red-shift (the
blue curve), with the upper axis showing the cosmic evolution time since the Big Bang
(using parameters from REF PLANCK). The bandwidth for HERA is demarcated by the
horizontal dashed lines. The black dashed lines denote the ``EOR'' core bandwidth,
and the extension to the blue dotted lines is to capture earlier epochs (going lower
in frequency) and find the post-EOR ``null'' (going higher in frequency) as discussed
below. Table \ref{tab:heraband} provides the details of how the frequencies map to
redshift and cosmic time.

\subsection{Bandwidth Field-of-View, and Spatial Scale}
As we've seen above, the frequency at which the hydrogen transition is observed
determines the source's cosmological distance. Therefore the bandwidth of the
observations defines a linear extent along the line-of-sight in space, specified by
the frequencies at the end points. As before, one may compute this extent as a
function of redshift using a cosmological calculator. Note that at large cosmological
distances, relatively narrow bandwidths can imply very large extents, which may have
significant cosmological evolution and hence are effectively different samples of the
Universe. For HERA, this bandwidth limit is about 10 MHz or less. The cosmological
extent for 10 MHz is shown in Figure \ref{fig:heraXY}.

Telescopes have a field-of-view of the observed sky, which gives the two orthogonal
directions to the line-of-sight. A given diameter and redshift will then have a
plane-of-sky extent that with the bandwidth will define a cosmological volume. This
plane-of-sky extent for a 14-meter dish is also shown in \ref{fig:heraXY}.

\subsection{Delay}
To exploit the smooth spectral response filtering of the foregrounds

\subsection{Sensitivity Metric}
The equation for sensitivity has been described in previous works [ref ] and is reproduced below:
\begin{equation}
\centering
\label{eq:sensitivity}
\Delta^2_N(k) 
\end{equation}
Note that $\Delta^2_N(k)$ is the standard radiometric sensitivity equation, scaled by
the volume in $k$-space, normalized by the power spectrum Fourier coefficient, and
reduced by the number of independent samples in a given $k$-mode bin, which may have
both coherent and incoherent application.

Why close-packed.  Hexagonal has 60degree symmetry, not 90degree symmetry like for rectangular.

Why 14-m dishes.

\section{System Overview}
\subsection{Antenna}
14-m

\subsection{Analog System}
amplifiers and stuff

\subsection{Digital System}
fpga and gpus and stuff

\section{Site}
Karoo

\conclusions  %% \conclusions[modified heading if necessary]
And sugar and spice

%\appendix
%\section{\\ \\ \hspace*{-7mm} HEADING}    %% Appendix A
%And everything nice.
%\subsection                               %% Appendix A1, A2, etc.
%And everything else.

\begin{acknowledgements}
TEXT
\end{acknowledgements}

\begin{thebibliography}{}

\bibitem[AUTHOR(YEAR)]{LABEL}
REFERENCE 1

\bibitem[AUTHOR(YEAR)]{LABEL}
REFERENCE 2

...

\end{thebibliography}

%% Literature citations
%% command                        & example result
%% \citet{jones90}|               & Jones et al.\ (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[ht]
\vspace*{2mm}
\begin{center}
\includegraphics[width=\textwidth]{plots/herauniall.png}
\end{center}
\caption{\small Cartoon showing the development of the Universe and HERA's observing window into it (dashed lines).  The blue solid
line shows the \HI transition frequency as a function of redshift/cosmic evolution age.  The black dashed lines indicate the HERA EOR ``core'' band and blue dashed lines the expanded band, which are detailed in Table \ref{tab:heraband}.}
\label{fig:theUniverse}
\end{figure*}

\begin{figure}[t]\centering
\includegraphics[width=\columnwidth]{plots/eor_pspec_2014.png}
\includegraphics[width=\columnwidth]{plots/hera_pk3pk.png}
\caption{\small Top: HERA's power-spectrum sensitivity (solid)
relative to a fiducial ionization model (dotted line; $\xHI=0.37$, $z=9.0$).  
Sensitivities reflect staged array size and
improving analysis software that expands the range
of modes free of foreground systematics. 
Bottom: The current best upper limit on the 21~cm reionization power spectrum,
obtained with a 32-element PAPER deployment \citep{parsons_et_al2013}.  These upper limits
constrain the brightness temperature of the IGM at $z\sim8$, showing
a departure from adiabatic cooling presumed to be indicative of X-ray heating.
}\label{fig:eor_pspec}
\end{figure}

\begin{figure}[t]
\vspace*{2mm}
\begin{center}
\includegraphics[width=\columnwidth]{plots/heraXY.png}
\end{center}
\caption{Scaling  \label{fig:heraXY}}
\end{figure}

%% TWO-COLUMN FIGURES

%f
%\begin{figure*}[t]
%\vspace*{2mm}
%\begin{center}
%\includegraphics[width=12cm]{FILE NAME}
%\end{center}
%\caption{TEXT}
%\end{figure*}

%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[t]
\caption{HERA band in frequency, redshift, cosmic age and comoving distance.}
\label{tab:heraband}
\vskip4mm
\centering
\begin{tabular}{| c | c | c | c |} \hline
\textbf{Frequency [MHz]} & \textbf{Redshift} & \textbf{Age} & \textbf{Look-back} \\
\textbf{[MHz]}                   &                            & \textbf{[Myr]}  & \textbf{[Gly]} \\ \hline
50 & 27.4    & 164 & 13.6 \\ \hline
100 & 13.2  & 372 & 13.4 \\ \hline
200  & 6.1    & 960  & 12.8 \\ \hline
225  &  5.3   & 1,135  &  12.7 \\ \hline
\end{tabular}
\end{table*}

%% ONE-COLUMN TABLE

%t
%\begin{table}[t]
%\caption{TEXT}
%\vskip4mm
%\centering
%\begin{tabular}{column = lcr}
%\tophline

%\middlehline

%\bottomhline
%\end{tabular}
%\end{table}

%% TWO-COLUMN TABLE

%t
%\begin{table*}[t]
%\caption{TEXT}
%\vskip4mm
%\centering
%\begin{tabular}{column = lcr}
%\tophline

%\middlehline

%\bottomhline
%\end{tabular}
%\end{table*}

%% The different columns must be seperated with a & command and should
%% end with \\ to identify the column brake.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% If figures and tables must be numbered 1a, 1b, etc. the following command
%% should be inserted before the begin{} command.

%\addtocounter{figure}{-1}\renewcommand{\thefigure}{\arabic{figure}a}

\end{document}

### 2014-05-07 hera-concept2013.tex
update...

@@
 %\documentclass[ars, hvmath, online]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus_discussions}
-
+\usepackage{amsmath}
 \def\kperp{k_{\bot}}
@@
 the Universe, which at that time existed as a nearly uniform sea of warm neutral
-hydrogen.   This period of cosmic reionization represents one of the few global
+hydrogen.  Figure \ref{fig:theUniverse} is a cartoon showing cosmic evolution from the Big Bang on the left through
+to HERA today. The period of cosmic reionization represents one of the few global
 phase changes in the evolution of the Universe.
@@
 (IGM), and cosmology.
-
-Figure \ref{fig:theUniverse} is a cartoon showing cosmic evolution from the Big Bang on the left through
-to HERA today.
 
@@
 
-HERA overall is a phased set of experiments that build on its predecessors and preceding stages.
+The HERA roadmap is a phased set of experiments that build on its predecessors and preceding stages.
 The first phase of the HERA roadmap entailed the operation of the
@@
 
-Building on this understanding, the next stage of HERA entails a new
+Building on this understanding, the next stage of HERA incorporates a new
 14m diameter antenna element that is optimized both for sensitivity
@@
 HERA telescope to be built to detect and characterize the EoR power
-spectrum.
-
-\section{Background}
-\label{sec:background}
-\subsection{Hubble's Law}
-\label{sec:hubble}
-Modern cosmology has provided a spectacular understanding of the cosmic history of
-our Universe starting from the instantaneous moment after the explosion of our
-spacetime that we call ``the Big Bang'', which occurred about 13.8 Gyr ago, to the
-advent of the large structures that seeded our cosmic current structure. After the
-Big Bang, matter and space went streaming outward and is manifest to us by the simple
-expression called Hubble's Law:
-\begin{equation}
-v = H_od
-\end{equation}
-where $v$ is the velocity of the object away from us, $d$ is the distance to the
-object $H_o$ is the so-called ``Hubble Constant'', which has a value of 68 Mpc/km/sec
-\ref{planck}, in astronomer's units, where Mpc denotes megaparsecs, where a parsec is
-$3.086 \times 10^13$ km. Using measured Doppler shifts one may measure $v$, so if one
-knows $H_o$ one may directly compute the distance to the object. Velocity is
-typically given as a ``redshift'', denoted as $z$, related to one another by the
-non-relativistic Doppler shift equation $z \sim v/c$, where $c$ is the speed of light
-in vacuo. And then by Hubble's Law, that redshift conveys a distance.
-
-Note that Hubble's ``Constant'' is a function of time and using $H_o$ denotes today's
-value. Determining the evolution of Hubble's constant is essentially the role of
-cosmological studies and using Hubble's law for very distant sources requires an
-integration with cosmological evolution encoded within it. The values here are
-derived from \ref{ned_evo}.
-
-\subsection{Power Spectrum}
-\label{sec:pspec}
-One tool that is typically employed in studying large scale parameters of the
-Universe, is to collapse the richness of its spatial variation to its spatial Fourier
-transform and compute a power spectrum of this. It is the parameters of this power
-spectrum that may be predicted by cosmology, rather than the essentially random
-spatial variation about it. It is analogous to the fact that one may get a few broad
-parameters of an FM signal by Fourier transforming its time series, regardless of whether it is
-Beethoven or M\"{o}tley Cr\"{u}e, which have very different values as a function of time.
-
-Cosmology typically encodes this data as power in spherical harmonics.  Epoch of
-Reionization studies typically use wavevectors, $\k$.  This 3-D vector may be decomposed
-into plane-of-sky values (${\bf \kperp}$) and a line-of-sight term (${\kpar}$).  The overall
-wavevector may then be written as $\k = {\bf \kperp} + \kpar{\bf z}$.  As a wavevector,
-the units of $\k$ are 1/Mpc (inverse length).  Note that although modern cosmology believes
-it has pegged the value of Hubble's constant, to hedge their bets, they will often factor out
-$h=H_o/100$ as an overall scale adjustment, such that the units are $h$/Mpc.  The units
-of the power spectrum magnitudes are given in volume-adjusted power-squared (mK$^2$), denoted
-$\Delta ^2$.
-
-\subsection{Foregrounds and Overall Signal Levels}
-\label{sec:foregrounds}
-At the frequencies of interest (50-250 MHz), the sky as seen by a radio telescope is dominated 
-by the synchrotron emission from free electrons spiralling around the magnetic fields that thread
-through our Galaxy.  The magnitude of this signal varies as a power law of the frequency, as well
-as position on the sky (hottest in the plane of the galaxy, cooler at the poles).  The variation spans
-from about 10,000K to about 100K in equivalent temperature.  There are a few localized sources
-in the sky that are even hotter to the eyes of a radio telescope.
-
-As will be shown shortly, the signals of interest have equivalent temperatures of $<$ 10 mK, 7 to 9 
-orders of magnitude below the foreground signal strengths.  The challenge is to peer through these
-strong foreground signals to see the weak signal of interest below.
-
-As stated above, the strong foreground signals may be accurately described as smooth power laws in 
-frequency - a factor that may be exploited to extract the signals of interest, which have a structure
-that varies fairly rapidly with frequency.
-
-It is also interesting to note that another ``foreground'' from our perspective is the Cosmic Microwave Background.
+spectrum.  After presenting the scientific and motivation and background, the
+derived specificiations and resulting system will be described.  The data pipeline
+will then be summarized, follwed by the conclusion.
 
@@
 
+\section{Background}
+\label{sec:background}
+\subsection{Hubble's Law}
+\label{sec:hubble}
+Modern cosmology has provided a spectacular understanding of the cosmic history of
+our Universe starting from the instantaneous moment after the explosion of our
+spacetime that we call ``the Big Bang'', which occurred about 13.8 Gyr ago, to the
+advent of the large structures that seeded our cosmic current structure. After the
+Big Bang, matter and space went streaming outward and is manifest to us by the simple
+expression called Hubble's Law:
+\begin{equation}
+v = H_od
+\end{equation}
+where $v$ is the velocity of the object away from us, $d$ is the distance to the
+object $H_o$ is the so-called ``Hubble Constant'', which has a value of 68 Mpc/km/sec
+\ref{planck}, in astronomer's units, where Mpc denotes megaparsecs, where a parsec is
+$3.086 \times 10^13$ km. Using measured Doppler shifts one may determine $v$, so if one
+knows $H_o$ one may directly compute the distance to the object. Velocity is
+typically given as a ``redshift'', denoted as $z$, related to one another by the
+non-relativistic Doppler shift equation $z \sim v/c$, where $c$ is the speed of light
+in vacuo. And then by Hubble's Law, that redshift conveys a distance.
+
+Note that Hubble's ``Constant'' is a function of time and using $H_o$ denotes today's
+value. Determining the evolution of Hubble's constant is essentially the role of
+cosmological studies and using Hubble's law for very distant sources requires an
+integration with cosmological evolution encoded within it. The values here are
+derived from \ref{ned_evo}.
+
+\subsection{Power Spectrum}
+\label{sec:pspec}
+One tool that is typically employed in studying large scale parameters of the
+Universe, is to collapse the richness of its spatial variation to its spatial Fourier
+transform and compute a power spectrum of this. It is the parameters of this power
+spectrum that may be predicted by cosmology, rather than the essentially random
+particular spatial variation about it. It is analogous to the fact that one may get a few broad
+parameters of an FM signal by Fourier transforming its time series, regardless of whether it is
+Beethoven or M\"{o}tley Cr\"{u}e, which have very different values as a function of time.
+
+Cosmology typically encodes this data as power in spherical harmonics.  Epoch of
+Reionization studies typically use wavevectors, $\k$.  This 3-D vector may be decomposed
+into plane-of-sky values (${\bf \kperp}$) and a line-of-sight term (${\kpar}$).  The overall
+wavevector may then be written as $\k = {\bf \kperp} + \kpar{\bf z}$.  As a wavevector,
+the units of $\k$ are 1/Mpc (inverse length).  Note that although modern cosmology believes
+it has pegged the value of Hubble's constant, to hedge their bets, they will often factor out
+$h=H_o/100$ as an overall scale adjustment, such that the wavevector ``units'' are $h$/Mpc.  The units
+of the power spectrum magnitudes are given in $\k$-volume-adjusted power-squared (mK$^2$), denoted
+$\Delta ^2$.
+
+Note that the initial approach to detecting and characterizing the EOR is therefore not imaging, but 
+to detect power in a few of the power spectrum bins.  One 
+may exploit this to get more samples of the power in specific bins to beat down the noise, at the expense of 
+imaging, although for very compact arrays there is less of an imaging penalty as more of the bins
+are filled within a radius of support.  The implications of the configuration to the array sensitivity
+will be touched upon below.
+
+\subsection{Foregrounds and Overall Signal Levels}
+\label{sec:foregrounds}
+At the frequencies of interest (50-250 MHz), the sky as seen by a radio telescope is dominated 
+by the synchrotron emission from free electrons spiralling around the magnetic fields that thread
+through our Galaxy.  The magnitude of this signal varies as a power law of the frequency, as well
+as position on the sky (hottest in the plane of the galaxy, cooler at the poles).  The variation spans
+from about 10,000K to about 100K in equivalent temperature.  There are a few localized sources
+in the sky that are even hotter to the eyes of a radio telescope.
+
+As will be shown shortly, the signals of interest have equivalent temperatures of $<$ 10 mK, 7 to 9 
+orders of magnitude below the foreground signal strengths.  The challenge is to peer through these
+strong foreground signals to see the weak signal of interest below.
+
+As stated above, the strong foreground signals may be accurately described as smooth power laws in 
+frequency - a factor that may be exploited to extract the signals of interest, which have a structure
+that varies fairly rapidly with frequency.
+
+It is also interesting to note that another ``foreground'' from our perspective is the Cosmic Microwave Background.  
+In this case, the detectable signal is the difference in flux between the hydrogren and the CMB.  It acts a bit like the
+weatherman's "green screen", where anything in a specfic shade of green is invisible.
+
+\subsection{Filtering}
+The wedge etc
+
 \section{Specifications}
@@
 in frequency) and find the post-EOR ``null'' (going higher in frequency) as discussed
-below. Table \ref{tab:heraband} provides the details of how the frequencies map to
-redshift and cosmic time.
+below. Figure \ref{fig:theUniverse} shows the mapping.
 
@@
 
-\subsection{Delay}
+\subsection{Delay and Systematics}
 To exploit the smooth spectral response filtering of the foregrounds
 
-\subsection{Sensitivity Metric}
-The equation for sensitivity has been described in previous works [ref ] and is reproduced below:
+\subsection{Sensitivity}
+The sensitivity specification is to maximize performance per cost, which determines the element 
+diameter ($D$) and the number of elements ($N$) subject to the constraints above.  One therefore 
+needs a model of cost and performance as a function of $D$ and $N$.  Given a fairly mature element
+and system design, a bottoms-up costing appropriate for element diameters from about 6-m to 20-m 
+has been done for hex-numbers corresponding to element counts from 37 to 631.  The costs here
+are only those associated with delivering the array to that scope on-site, so excluding development and
+science.
+
+The equation for sensitivity has been described in previous works [REFS] and depends on many parameters
+related to the instrumentation, the configuration, the location, the observing strategy, etc.  A useful form for 
+the proposed compact configuration is Eq 25 in Parsons et al 2012 and reproduced here:
 \begin{equation}
@@
 \label{eq:sensitivity}
-\Delta^2_N(k) 
+\begin{split}
+\Delta^2_N(k) \approx 60 \left[\frac{k}{0.1h\text{Mpc}^{-1}}\right]^\frac{5}{2}
+                                         \left[\frac{6 \text{MHz}}{B}\right]^\frac{1}{2}
+                                         \left[\frac{1}{\Delta\ln k}\right]^\frac{1}{2} \\
+                        \times       \left[\frac{\Omega}{0.76\text{sr}}\right]
+                                         \left[\frac{T_\text{sys}}{500 \text{K}}\right]^2
+                                         \left[\frac{6 \text{hrs}}{t_\text{per\_day}}\right]^\frac{1}{2} \\
+                        \times       \left[\frac{120\text{days}}{t}\right]
+                                         \left[\frac{32}{N}\right]
+                                         \left[\frac{\text{10}^4f_o}{f}\right]  \text{mK}^2
+\end{split}
 \end{equation}
+where $k$ is the magnitude of the $k$-mode, $B$ is the bandwidth, $\Delta\ln k$ is the log
+of the binsize, $\Omega$ is the field-of-view, $T_{\text{sys}}$ is the system temperature, 
+${t_\text{per\_day}}$ is the number of hours observed per day, $t$ is the number of days
+observed, $N$ is the number of antennas, and $f_o/f$ is the configuration metric for a 
+redundant array as defined in Parsons et al.
+
 Note that $\Delta^2_N(k)$ is the standard radiometric sensitivity equation, scaled by
@@
 
-Why close-packed.  Hexagonal has 60degree symmetry, not 90degree symmetry like for rectangular.
+Pulling out terms relating to diameter and number, we can write Eq. \ref{eq:sensitivity} as
+\begin{equation}
+\label{eq:reducedSensitivity}
+\Delta^2_N(k) \propto \frac{\Omega (f_o/f)}{N\sqrt{t_\text{per\_day}}} \propto \frac{(1/D^2)(1/\sqrt{N})}{N\sqrt{D}}
+= D^{-\frac{5}{2}}N^{-\frac{3}{2}}
+\end{equation}
+where the dependencies on diameter and number have been subsituted in, noting that the expressions for $f_o/f$ and 
+$t_\text{per\_day}$were derived in Parsons et al where the baselines for the close-packed array are multiples of the diameter.
 
-Why 14-m dishes.
+For a fixed sensitivity, we may then parameterize the performance/cost ratio by the diameter.  Figure \ref{fig:costing}
 
 \section{System Overview}
-\subsection{Antenna}
-14-m
+\label{sec:design}
 
-\subsection{Analog System}
-amplifiers and stuff
+We have achieved a
+pivotal new understanding of how instrumental characteristics give rise to the
+wedge of emission shown in Figure \ref{fig:twoFGViews}.  Furthermore, we have
+measurements that prove the efficacy, to the sensitivity limits of current
+instruments, of projecting out foreground-dominated wavemodes within the
+wedge.  The HI cosmology community is now in a position to define
+the top-level instrument requirements that ensure foregrounds remain bounded
+within the wedge, and to specify the sensitivity needed to obtain
+high-significance detections of the 21~cm reionization power spectrum under the
+conservative assumption that all foreground-dominated wavemodes within the
+wedge must be projected out of our measurements.
+As summarized in Table \ref{tab:signif} (see \Mycitealt{pober_et_al2014} for more details),
+based on these assumptions, current instruments are likely to achieve,
+at best, only marginal detections of power from reionization.
+
+%HERA follows the vision delineated in NWNH where the MWA and PAPER
+%characterized the foreground emission and performed the first deep
+%integrations, followed by a more powerful instrument that drew on the lessons
+%learned and merged the scientific teams. The MWA, PAPER, and LOFAR have all
+%recently started deep integrations with hopes of detecting the 21~cm
+%signal.
+%However, as summarized in Table \ref{tab:signif} (see \citealt{pober_et_al2014} for more details),
+%these instruments are likely to achieve
+%at best only marginal detections of power from reionization.
+%While this might qualify as a ``detection'' if the signal is strong, the
+%\emph{science} of reionization observations is beyond their reach. 
+%However, the profound contributions of MWA and PAPER should not be overlooked.
+PAPER and the MWA have succeeded magnificently in characterizing the foreground emission, discovering the EoR window, and understanding how to remove foreground contamination from the measurements.
+Further, they have developed and retired the risk associated with all of HERA's major systems.
+%HERA uses these lessons perform the science of reionization, at a much lower cost than assumed in NWNH.
+%% that this proposal is cheaper than NWNH suggested but never give either that
+%% number nor your estimated cost.
+
+This proposal targets two arrays: a 127-element array that borrows heavily
+from components of the PAPER experiment, and an upgraded 331-element array that
+incorporates several performance optimizations.  
+These arrays are developed over four years (see Fig. \ref{fig:timeline} and the Project Management Plan for details),
+with recurring cycles of development, testing, review, deployment, commissioning, and observation.
+As listed in Table \ref{tab:signif}, the specifications of HERA-127
+and HERA-331 have been set to meet precisely the requirements for obtaining,
+respectively, a 10$\sigma$ detection of the 21~cm reionization signal across a broad range of redshifts, and a
+25$\sigma$ detection of the power spectrum of fluctuations in the 21~cm signal
+capable of determining the nature and distribution of the first galaxies that
+dominate cosmic reionization.  
+As discussed in
+\S\ref{sec:Lessons}, these science requirements translate directly into
+requirements for signal delay and collecting area that, when 
+combined with a cost minimization requirement, tightly
+bound the HERA design.  The basic parameters are given in Table \ref{tab:BasicParameters}.
+
+%incorporates
+%our proven foreground avoidance techniques while improving
+%dramatically the sensitivity relative to current experiments.  With a
+%new understanding of how antenna size and separation affect
+%sensitivity and foreground isolation, it has become evident a revision
+%of the PAPER antenna design can yield up to 20 times the sensitivity
+%per element without substantially degrading foreground isolation.
+%Where PAPER's elements lack collecting area and are smaller than
+%strictly required for foreground isolation, and the majority of MWA
+%and LOFAR elements are spaced too widely to avoid foregrounds,
+%HERA-331 employs an extremely compact array of 14-m parabolic dishes
+%with PAPER-style dipole feeds (see Figure \ref{fig:hera_dish}).  The dishes
+%are placed in a compact hexagonal configuration of 331 dishes supplemented by 18 outrigger dishes to achieve dense and
+%highly-redundant $uv$-sampling (see Figures \ref{fig:uv_coverage} and \ref{fig:config_optics}).  The dishes have a short (4.5m) focal height to limit the
+%path length of reflections, whose time-delay gives rise to chromatic
+%instrumental systematics.
+%
+%
+%% AAARRGGHH!
+%%\begin{figure}[t]
+%%\centering
+%%	\begin{subfigure}[b]{0.46\textwidth}
+%%		\includegraphics[width=\textwidth]{plots/paper_element.jpg}
+%%		\Caption{-0.2in}{0.99}{-0.2in}{The PAPER element (provides a clean instrumental response as a function
+%%		of frequency \citep{parsons_et_al2010,parsons_et_al2012b}, which is crucial to
+%%		the foreground isolation shown in Figure \ref{fig:eor_pspec}.}
+%%	\end{subfigure}
+%%	\quad
+%%	\begin{subfigure}[b]{0.46\textwidth}
+%%		\includegraphics[height=1.75in]{plots/hera_dish.png}
+%%		\Caption{-0.2in}{0.99}{-0.2in}{A 14m dish designed around the feed dramatically improves sensitivity while
+%%		constraining the path length and amplitude of reflections to ensure that foreground 
+%%		isolation is not substantially degraded.}
+%%	\end{subfigure}
+%%\Caption{-0.2in}{0.99}{-0.2in}{PAPER and HERA elements}
+%%\label{fig:hera_dish}
+%%\end{figure}
+%
+%The size of HERA-331 dishes optimizes cost for a fixed sensitivity and
+%level of foreground isolation.  The associated reduction in the number
+%of antenna elements to achieve a given collecting area, combined with
+%the fact that these dishes have no moving parts, are built from
+%inexpensive materials, and follow a simple construction that can be
+%contracted locally, makes the cost of building HERA-331 substantially
+%cheaper than was anticipated in the roadmap submitted to \nwnh\ for this
+%stage of the program.   
+%
+%HERA leverages the technical heritage of PAPER, MWA and of CASPER\footnote{Collaboration for Astronomical
+%Signal Processing and Electronics Research, a Berkeley-initiated worldwide open source community that is
+%developing boards, firmware and software for the astronomical community} and incorporates a
+%phased implementation to mitigate against risk.  The system, technology and phased approach is discussed below
+%in more detail.
+%The HERA instrument design is very straightforward.  The element
+%itself is a fixed zenith-pointing 14-meter segmented prime-focus paraboloid with a high screen
+%to minimize cross-talk between elements (which are spaced 14.3 meters on a
+%hexagonal grid).   The $f/D$ of
+%the paraboloid is 0.32, so that the focal length, $f$, is less than 5 meters to meet the 
+%standing wave specification at the delays of interest of more than 60 dB of attenuation at delays 
+%greater than 50 ns.
+%
+%
+%The active feed sends back the entire dual-polarization analog bandwidth on standard
+%coaxial cable to an aggregation point called a ``node'', which services 
+%about 15 antennas.  This cable length is kept short (35-m) to keep any standing
+%wave contamination outside of the delay-space of interest for power spectrum
+%measurements.  The node amplifies, filters, digitizes and transmits the signal data stream
+%back to the central processing location (the Karoo Array Processing Building - KAPB).
+%Figure \ref{fig:blockDiagram} shows a block diagram of the system.
+
+As shown in Figure \ref{fig:systemOverview}, the HERA instrument has a straightforward signal
+path from antenna elements
+with active feeds, to nodes that digitize and aggregate signals onto an optical network,
+and on to a processing building where signals are correlated and data are stored, compressed,
+and shipped to a computing cluster located at U. Pennsylvania (UPenn).
+
+\subsection{Antenna Element}
+
+The novel design of HERA's antenna element
+(Fig. \ref{fig:hera_dish}, right panel) represents a critical advance
+that enables HERA to achieve its science goals.  This 14-m
+fixed zenith-pointing parabolic dish yields more than 10 times the
+sensitivity of an MWA tile (and more than 20 times
+that of a PAPER element), but does so without substantially degrading
+our ability to isolate and remove foreground emission on the basis of
+spectral smoothness.  As described above, this is done by limiting
+the timescale of signal delays and reflections to under 120 ns.
+
+Since the time it takes radio emission to propagate between adjacent antennas 
+represents a significant portion of this time budget, HERA elements must be placed close together.
+Recent work characterizing foregrounds suggests that 
+antenna separations of $8\lambda \approx 15$m are
+well-behaved to current limits \Mycitep{pober_et_al2013,parsons_et_al2013}. This influences our
+choice of a 14-m dish diameter,
+which incurs 42 ns of signal delay between adjacent antennas.
+The focal height (4.5 m) derives from the illumination pattern of the prime-focus
+feed and the fact that reflections
+between the feed and the element
+%caused by imperfect impedance matching between the feed electronics and free space, 
+can introduce additional signal delay. 
+%With a splash cone underneath the feed to scatter reflections, we assume -20 dB
+%attenuation per reflection, so that after three traversals (13.5 m; 41 ns), reflections from
+%foreground emission are below the expected level of the 21~cm reionization signal.
+Additional measures, such as a splash cone underneath the feed, screens that isolate feeds from one another, and 
+non-metallic cabling and poles for supporting the feeds, are all aimed at minimizing
+the potential for additional sources of signal reflections.
+The total signal delay associated with the HERA dish design is estimated at 83 ns, leaving
+headroom for reflections arising in the analog signal path.
+
+Cost is another design constraint, including the price of construction
+materials, assembly in a remote area, and maintenance over the operational
+lifetime of the array.  Care has been taken to select robust and inexpensive
+construction materials (PVC, concrete, utility poles, 0.25" wire cloth) and an assembly methodology that delivers the required positional
+accuracy of 10 cm and surface accuracy of 2 cm given standard expertise in construction
+practices for the subcontracted teams that set the poles and construct the
+elements in the field.  
+%JCP: "required positional accuracy" -- what drives the requirement?
+Given the sensitivity and signal delay requirements, the size of HERA
+dishes optimizes a global costing curve that includes the costs of the elements,
+the signal path, correlation, data storage, and processing.
+
+% ARP: removing this figure, since it appears in project management plan
+%\begin{figure}[h]
+%	\centering
+%		\includegraphics[width=0.46\textwidth]{plots/heracles.png}
+%		\includegraphics[width=0.46\textwidth]{plots/Engineering/heraclesNA.png}
+%	\caption{\small
+%		The construction prototype element in California (left) with preliminary 
+%        reflection measurements (right).
+%} \label{fig:heracles}
+%\end{figure}
+
+%As illustrated in Figure \ref{fig:heracles}, 
+A first prototype of the HERA dish is nearly completely constructed.  In this
+proposal, UC Berkeley and NRAO 
+lead the incorporation of lessons from the
+construction process and reflectometry tests into the design.
+Two further prototype elements are constructed in the first year alongside the existing
+PAPER array in Green Bank, WV,
+for on-sky measurements of the beam pattern and spectral 
+response.
+These measurements are coupled with full electromagnetic modeling to test
+and refine the element design prior to a Critical Design Review at the end of the first project
+year.  Three revised elements are then constructed in South Africa, involving
+lead members of the construction subcontract teams for a Project Readiness Review.  %Thereafter,
+%element construction will proceed continously through the second and third project years during the day,
+%with a staged incorporation of new elements for night-time observing.
+
+%The elements are
+%also nearly abutting one another, which allows for sharing of physical support
+%infrastructure.
+
+% ARP: lack room for this figure.
+%\begin{figure}[h]
+%	\centering
+%	\begin{subfigure}[b]{0.46\textwidth}
+%		\includegraphics[width=0.99\textwidth]{plots/Engineering/nvsd.png}
+%		\caption{Costing model for a fixed sensitivity and varying the diameter.   The arrows indicate the regions of 
+%				increasing systematics and delay-space contamination.}
+%		\label{fig:nvsd} 
+%	\end{subfigure}
+%	\quad
+%	\begin{subfigure}[b]{0.46\textwidth}
+%		\includegraphics[width=0.8\textwidth]{plots/Engineering/focalEff.png}
+%		\caption{Analytical model efficiency of a parabolic element as a function of focal height and diameter.   The delay 
+%				contamination specification is $f<5$m.}
+%		\label{fig:disheffic}
+%	\end{subfigure}
+%	\caption{Cost and performance variational analysis data for the element.}
+%\end{figure}
+
+% FOLLOWING SEEMS TOO DETAILED FOR A PROPOSAL
+%To distinguish between measured sky delays and instrumental-induced delays at a required threshold ($R_T$) dB, 
+%we need sufficient attenuation of the reflected signals at a given focal length ($f$) at a specified delay ($\tau_{d}$).
+Assuming a conservative simple model that the magnitude of the reflection is attenuated by $A$ dB at each reflection
+%we see that for a delay length limit of $\delta_{d}=c\tau_{d}$ and focal length $f$,  the required focal length is
+%
+%\begin{equation}
+%f < \left(\frac{A}{R_T}\right)\delta_d.
+%\end{equation}
+%
+%Using nominal values of $R_T$ =  60dB (an order of
+%magnitude below where EoR is predicted to be below foregrounds) at delays
+%corresponding to the time it takes travel 15m and a net attenuation of 20 dB per reflection, 
+%we find that the focal length should be less than about 5 m.  Free-space loss effects would 
+%increase that value, loosening the constraint.
+%DETAILEND
+
+%In addition to the constraints given by the element itself, the HERA element size is
+%also influenced by the location of the ``knee'' in the EoR power spectrum \citep{lidz_et_al2008}
+%The EoR power spectrum has an upward slope for low $k$-modes which levels off
+%around $k=0.15 h$/Mpc %(see fig blah). 
+%Working inside this $k$-mode would be beneficial due to the fact foregrounds are
+%less problematic. This poses a problem because without knowing the
+%width of our foregrounds, we can't say for sure which $k$-modes (in the power
+%spectrum) are corrupted. This uncertainty, coupled with increasing systematic affects for shorter 
+%baselines (hence smaller diameters), favors larger diameter antennas.
+
+%% ARP: no room for this figure.
+%\begin{figure}[h]
+%	\centering
+%	\begin{subfigure}[b]{0.46\textwidth}
+%		\includegraphics[width=0.85\textwidth]{plots/dish.png}
+%		\caption{CAD model of 14m dish with screening and some supports removed to show detail.}
+%		\label{fig:dish} 
+%	\end{subfigure}
+%\quad
+%	\begin{subfigure}[b]{0.46\textwidth}
+%		\includegraphics[width=\textwidth]{plots/Engineering/hera_beam.png}
+%		\caption{Analytical model beam patterns at 10m, 12m and 14m.}
+%		\label{fig:beam} 
+%	\end{subfigure}
+%	\caption{HERA element and beam pattern.}
+%\end{figure}
+
+%The $k$-mode for a $14m$ baseline is $k = 0.023$, given by $k_{H} =
+%\frac{B}{c}\frac{dk}{d\eta}$, where $B$ is the length of the baseline, 14m in
+%our case, $c$ is the speed of light, and $\frac{dk}{d\eta}$ is the cosmological
+%transfer function from delays to $k$-modes. In addition, narcissistic
+%reflections add into our $k$ budget as well. The $f$ and $D$ choices above provide a $k$ budget for
+%foreground widths to be within $\Delta{k}\sim{0.1}$. Testing these hypothesis
+%and again, finding a compromise between maximum sensitivity, foreground budget
+%will be key to testing and constructing the required element. Foreground width constraints
+%are still an area of active research \citep{pober_et_al2013}.
+
+% The core defining elements in this design are the central hub and three tall support poles.  Three intermediate 
+% support posts are installed between each pair of poles.  A 2$^{\prime\prime}$ PVC spar of 24.1$^{\prime}$ 
+% terminates at each pole and post.  These spars are supported at each end and one point in the middle at the 
+% proper height and angle.  The intervening PVC pipe essentially acts as a smoothing filter between those 
+% points, noting also that a beam with point loads attains nearly the quadratic shape desired.  The CAD model is 
+% shown in the right panel of Figure \ref{fig:hera_dish}.  
+% 
+% The tall ($\sim$ 7m) poles provide locational accuracy (in all three dimensions) for the overall array installation.  
+% Using conventional commercial pole-installation techniques the poles are installed first for the entire array.  
+% Note that every pole except for the edge poles are shared by three antennas.  A standard theodolite can 
+% then be used to mark a known level height on all three poles.  These locations are then used with tensioned 
+% lines to define the center of that element and the hub is positioned at that location using a jig.
+% 
+% The hub uses concentric commercially available ``sonotube'' forms (circular cardboard forms for concrete pillars) 
+% and PVC sleeves to hold the PVC spars and PVC supports.  The retaining holes may be accurately cut into 
+% the forms, sleeves installed and concrete poured to make a simple hub to the desired accuracy.  The jig holds 
+% the concentric rings in place and allows it to be centered by tensioned lines while the concrete is poured.  
+% When the concrete cures one can then transfer an accurate offset from the dish vertex back to the poles.
+% The intermediate posts are then located by the support sub-assemblies attached to the poles and posts along 
+% with the rim sub-assemblies.  These are positioned and a small pier is poured to locate them.  After spar support 
+% pieces are installed on the posts, the spars themselves and the metal cloth can be installed.
+% 
+% The feed is held off the three tall poles using tensioned lines to accurately locate it over the hub.  A precise 
+% length of kevlar rope holds the feed down to the hub at a precise focal point.  The RF cables follow the line 
+% down and out to the analog-to-digital converters and correlator.  
+% 
+% To minimize cross-talk, metal screens are strung between every pole/post, which go to the level of the feed.  
+% The dishes have a rim-to-rim spacing of 30cm to allow the screen to be slightly angled to minimize standing waves.
+
+\subsection{Array Configuration}
+
+ Figure \ref{fig:config}
+
+For HERA-127 and HERA-331, elements are arranged in a compact hexagonal grid.
+This configuration minimizes antenna separation, which is critical for meeting
+the science requirements for foreground isolation, and 
+produces a highly redundant sampling of the $uv$ plane.  As described in
+\S\ref{sec:Lessons}, redundant configurations have been employed by PAPER to
+boost sensitivity \Mycitep{parsons_et_al2012a}, with the added benefit that they
+greatly facilitate fast and accurate calibration
+\Mycitep{liu_et_al2010,parsons_et_al2013}.  Placing HERA elements in a
+grid allows certain construction components to be re-used between antennas,
+reducing cost and improving the accuracy of element placement.  HERA-331 
+is a build-out of HERA-127 from the eastern edge.
+
+The HERA-331 core has excellent
+imaging capability that can be leveraged for developing foreground suppression techniques
+to improve access to the 21~cm reionization signal 
+(Fig. \ref{fig:eor_pspec}, black).
+This capability is augmented with 21 outrigger elements.  These outriggers
+combine with the dense core to generate a fully
+sampled $uv$ plane out to 400$\lambda$.  While outriggers on the southern half of the array are
+placed in alignment with the hexagonal grid in the core, elements on the northern side are
+shifted off-grid to sample the $uv$ plane at sub-aperture scales.
+This sampling strategy helps eliminate grating lobes in the
+synthesized beam and provides information for
+calibrating and correcting direction-dependent antenna responses.  This capability may
+be used to match polarization beams and minimize
+polarization leakage.
+
+\subsection{Analog Signal Path}
+
+HERA's analog signal path inherits directly from PAPER.
+In keeping with a philosophy of
+incremental change,
+HERA-127 reuses the tested and commissioned feed, LNA, cables, and post-amplifier gain modules
+(Fig. \ref{fig:components}, left and center) that are currently being
+used to good effect in PAPER-128 \Mycitep{parsons_et_al2010}.  Existing PAPER feeds are attached to new
+reflector screens that are suspended over the elements, but otherwise remain unmodified.
+A parallel development effort in NRAO aims to improve the feed response below
+100 MHz 
+and to improve the match between polarization beams in the feed-element system.  
+Additional
+minor modifications to the analog signal path include transitioning from a 75$\Omega$ to a
+shorter 50$\Omega$ cabling system and replacing 100--200 MHz bandpass filters with 50--250 MHz equivalents.
+This wider analog bandwidth is used in HERA analysis to improve the extrapolation of smooth foreground
+emission over a broader range, and to explore HERA's capability as a Dark-Ages science instrument.
+After a Critical Design Review, the results of these development efforts are incorporated in
+the transition to the HERA-331 system.
 
 \subsection{Digital System}
-fpga and gpus and stuff
+\label{sec:digital}
 
-\section{Site}
-Karoo
+\begin{figure}[t]\centering
+\includegraphics[width=6.5in]{plots/casper_boards.jpg}
+\Caption{-0.3in}{0.99}{-0.15in}{\small
+Five generations of CASPER technology (progressing left to right) have been used to rapidly
+develop, test, and deploy digital instrumentation for radio astronomy.  This technology
+allows the PAPER correlator to be 
+easily upgraded for HERA, incorporating new technology
+\Mycitep{parsons_et_al2006,parsons_et_al2008}.
+}\label{fig:casper_boards}
+\end{figure}
+
+%HERA's digital system continues in the vein of PAPER's digital correlators.
+Although correlators have historically been one of the most complex,
+expensive, and risky aspects of developing a radio interferometer, this is no longer the case.
+CASPER \Mycitep{parsons_et_al2006}
+open-sourced the development of digital signal processing engines for astronomy and
+now has world-wide participation,
+with over 500 members at 73 institutions, and 
+five generations of hardware (Fig. \ref{fig:casper_boards}). 
+On a modest budget, PAPER has applied CASPER technology to develop and deploy new correlators
+annually for five years running, each quadrupling the computational capacity of its predecessor.
+Led at UC Berkeley's Radio Astronomy Lab (RAL),
+HERA efforts continue this incremental development cycle, following a packet-switched
+correlator architecture \Mycitep{parsons_et_al2008} that has been
+extended in recent PAPER and LEDA deployments (Fig. \ref{fig:components}, right)
+to leverage the computing strengths of both FPGAs and GPUs \citep{clark_et_al2011}.
+
+While HERA-127 uses the existing PAPER correlator directly, this correlator architecture evolves for
+HERA-331.  As discussed in previous sections,
+HERA's science requirements dictate a maximum analog signal path-length.  As a consequence,
+digitization needs to happen close to the antenna
+elements in the field.  This specification, along with a growing need for modularity to scale with the number of parallel signal paths,
+leads HERA to adopt a node-based architecture for amplification, digitization, channelization, and digital
+transmission in the field that builds on HERA's MWA heritage.  This architecture is merged with PAPER's clean 
+architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
+10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn that
+FX correlator architectures require.
+
+%\begin{figure}[t]\centering
+%%\includegraphics[width=6.5in]{plots/node_arch.png}
+%\caption{
+% something about nodes here.  Maybe also block diagram of NRAO digitizer board.
+%}\label{fig:node_arch}
+%\end{figure}
+
+{\bf Node}. HERA-331 employs RFI-tight node enclosures that each contain the final gain and digitization stages for
+signals from 18 antennas, along with power supplies, cooling, and a small server for monitor/control.  
+As part of HERA development,
+a new board called the Smart Network ADC Processor (SNAP) board is incorporated 
+into the CASPER suite of hardware and firmware. This inexpensive board was co-designed by UC Berkeley and NRAO to be
+both the digitizer and F-engine in HERA's FX correlator architecture,
+and is currently in layout at NRAO.  Each SNAP board 
+digitizes and channelizes a 50--250 MHz band for 6 input signals (3 antennas, dual-polarization).
+A 100-MHz band of selectable channels is transmitted over optical fiber
+to a central container (see below), and on to the Karoo Array Processing Building (KAPB).  The development and integration of the SNAP
+design and the node system is led at UC Berkeley and after a Critical Design Review, is fabricated
+under subcontracts to industry partners in the third project year and deployed as part of the HERA-331 system.
+Development activities under this proposal include porting the CASPER toolflow to
+the SNAP board, designing and testing the FPGA firmware,
+integrating and testing all components in the node subsystem, and providing a monitor/control
+access interface.  If time allows, optional development may include doubling the transmitted bandwidth.
+
+{\bf Central Container}.
+HERA's central container houses two significant subsystems adjacent to the array.  The first is a timing subsystem
+that maintains a GPS-disciplined oscillator and distributes timing
+signals (the sampling clock and 1PPS synchronization) to the nodes.  The second
+subsystem is a passive fiber optic patch panel that couples
+the optical network from the nodes into the 192-filament optical fiber bundle 
+that connects to the KAPB.
+
+{\bf Karoo Array Processing Building (KAPB)}.
+The KAPB is currently
+in advanced stages of construction for MeerKAT, and houses the switch and processors
+that 
+complete the HERA correlator system.  The fiber optic bundle that enters the KAPB patches
+into local fiber optic cables 
+that each terminate in optical transceivers that plug into a 240-port 10 GbE switch.
+Such switches, while large, are readily available commercially today.  Also connected to
+this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
+10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
+of the correlator during observations.  This estimate for the number of X-Engine servers
+is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
+speeds for transferring data into the GPU cores, but assuming that the computational
+capacity of such GPU cores doubles according to Moore's Law prior to the purchase of
+these servers in the third year of the project.
+Output data from the correlator are written to the data storage system described
+in the following section.
+
+%C. Post correlator data path
+% Aguirre, Moore
+
+\subsection{Data Storage, Compression, Transfer, and Computing}
+\label{sec:data}
+
+HERA's data management system is responsible for recording raw data from the
+correlator, compressing that data in real-time, applying routine calibration
+and analysis pipelines, and transferring data products to 
+a high-performance computing cluster at UPenn.
+UPenn leads the procurement
+and deployment of three data storage systems:
+\begin{itemize}[noitemsep,nolistsep]
+\item a 1.5 PB system deployed in the KAPB for archiving all 1.2 PB of raw data and 60 TB of compressed data, 
+\item 6 network attached storage (NAS) units plus 2 125 TB RAID array units that are used to ship compressed data products to the US, and
+\item a permanent 250 TB system at UPenn associated with a computing cluster that
+holds compressed data products and serves as the analysis engine for HERA collaborators.
+\end{itemize}
+\noindent
+This effort leverages existing infrastructure at UPenn, with support
+for the expansion of the data storage and for upgrading to a 30-node computing cluster.
+This cluster supports the bulk of the analysis by HERA collaborators (see \S\ref{sec:analysis}) that requires access to the full set of HERA observations.
+
+%The main data center at Penn will house all of the maximally compressed data from all seasons of HERA, as well as the LST-averaged data.
+The storage system is sized so that there is a factor of 4 overhead for work on the maximally-compressed data set, and a factor of 2 on the LST-averaged.
+The 500 TB data storage will be coupled to a 32-node cluster which performs the tasks of calibration and averaging to further reduce the data volume.
+This cluster will support the bulk of the analysis requiring the full data set.
+Subsets will be served off to collaborators from this system.
+The entire data storage and transfer plan for HERA draws heavily on the experience with PAPER data.
+
+The data compression scheme at the heart of the data management system
+has been implemented for PAPER (\Mycitealt{parsons_et_al2013},
+Appendix A), and is applied to HERA visibilities to reduce data volume by
+a factor of $\sim$20 without impacting reionization science
+capabilities.  This compression technique, which is based on delay/delay-rate
+filtering \Mycitep{parsons_backer2009}, is applied uniformly to all visibilities
+in the array, does not require (or produce) detailed calibration
+information, and is minimally restrictive for how data are analyzed and calibrated afterward.
+Data compression is run on the same GPU servers
+that implement the correlator X-Engines (\S\ref{sec:digital}).  Since HERA only observes at night,
+these processors would otherwise be unused.  UC Berkeley is responsible for porting
+the existing data compression pipeline to target these servers.
+
+Data quality assurance (QA) is performed in real time on an additional modest 
+computing cluster in the
+KAPB.  UPenn leads the deployment and support of the hardware system that
+manages data transfer, applies routine calibration pipelines (see \S\ref{sec:analysis}) and quality
+checks, and aggregates correlation-based metrics of array performance in real-time.  The QA system
+furnishes this information into the separate monitor and control system (\S\ref{sec:monitor}).
+A modest amount of data transfer is possible over the internet; PAPER's data transfer rate from the Karoo
+to UPenn varies, but peaks around 40 Mbps.  The QA system drives internet data transfer, but
+NAS devices and RAID storage, transferred by air-freight shipping, ensure full data transfer to UPenn.
+
+%ii. transport to data centers
+%In order to transfer maximally-compressed data over the internet in a timely manner, a network speed of around the instantaneous compressed data rate is required.
+Negotiations with the IT SKASA and the South African internet provider (Tenet) have achieved transfer speeds between the PAPER site and the United States have been measured to be up to 40 Mbps, which easily
+%accommodates the first two seasons of HERA observations.  In all seasons, however, portable, inexpensive, network-attached storage devices will be used to transfer data by air-freight shipping.
+
+%iii. data centers: access?
+
+%The main data center at Penn will house all of the maximally compressed data from all seasons of HERA, as well as the LST-averaged data.
+The storage system is sized so that there is a factor of 4 overhead for work on the maximally-compressed data set, and a factor of 2 on the LST-averaged.
+The 500 TB data storage will be coupled to a 32-node cluster which performs the tasks of calibration and averaging to further reduce the data volume.
+This cluster will support the bulk of the analysis requiring the full data set.
+Subsets will be served off to collaborators from this system.
+The entire data storage and transfer plan for HERA draws heavily on the experience with PAPER data.
+
+\subsection{Monitor and Control}
+\label{sec:monitor}
+
+The U. of Washington team leads the development of HERA's Monitor and Control (M\&C) system,
+which is a straightforward port of a similar system used on the MWA \Mycitep{tingay_et_al2013}.
+This system is
+responsible for tracking observing status, array startup and shutdown, and
+monitoring all active HERA systems. In the process, the M\&C system builds a sizeable database of 
+metadata that is crucial for verifying system functionality, identifying hardware failures, and feeding
+calibration and contextual information into data analysis pipelines.
+
+%Top-level interface for controlling and monitoring major subsystems. 
+%Collection of monitor data from nodes, switch, correlator, data compression, and data transfer subsystems.
+%Incorporation of calibration data from real-time application of redundant calibration (cite section), as well as
+%other calibration, imaging, and assessment subsystems.
+%
+%%\subsubsection{Array monitoring/maintenance: daily, weekly health monitoring}
+%% de Boer
+%
+%In addition to Monitor and Control subsystem, which aggregates routine checks that are run automatically as
+%part of the data compression and transfer pipeline, there is a need for deeper analysis
+%exploring the data for more subtle defects, as well as monitoring progress toward sensitivity goals.
+%We have set aside time for students at each institution, rotating among institutions, and managed at MIT,
+%for performing such data exercises.
+%Combine this with information from monitor database to identify hardware and subsystem failures.
+%Aggregate lists of systems requiring maintenance attention, which are then handled by system owners
+%in coordination with appropriate members of their teams.
+
+\subsection{Software, Analysis, and Science}
+\label{sec:analysis}
+
+Beyond the construction and data-taking aspects of HERA, this proposal
+incorporates a full data analysis effort, culminating in the publication of a
+suite of science papers connecting observations to the physics of cosmic dawn.
+These efforts leverage existing software pipelines, with on-going 
+development driven by students and postdocs targeting specific science goals.
+On the science side, HERA leverages the involvement of a team of theory collaborators,
+including Furlanetto, Lidz, Loeb, McQuinn, Mesinger, Oh, Pritchard, Santos, and Sutter.
+
+%\compress
+%\subsubsection{Calibration}
+%\label{sec:calibration}
+
+{\bf Calibration and Snapshot Imaging}. MIT leads
+the development of a real-time redundancy-based calibration pipeline based on related
+MITEoR and PAPER efforts.
+These instantaneous calibration solutions are provided to
+the Monitor and Control system to enable 
+hardware misbehaviors to be quickly identified, and also support the
+real-time imaging pipeline.  Absolute calibration is fixed with
+a combination of self-calibration techniques and absolutely calibrated baluns developed at ASU and NRAO.
+%Once data quality is assured, multi-day observations 
+%binned in local sidereal time, and averaged. This intermediate
+%averaging scheme will also generate an accurate visibility-based sky
+%model for the instrument once calibration is applied.
+Offline, detailed imaging efforts feed into developing
+empirical beam models that
+complement electromagnetic simulations.  UPenn develops polarization beam models
+that determine whether polarization leakage can be formally retired as 
+a risk and lay the foundation for imaging-based foreground suppression that
+corrects any remaining leakage effects.
+
+%\compress
+%\subsubsection{Foreground Modeling and Removal}
+%\label{sec:DataProducts}
+
+{\bf Foreground Modeling}. The U. Washington team leads the adaptation
+of a high dynamic-range imaging pipeline
+based on the Fast Holographic Deconvolution (FHD; \Mycitealt{sullivan_et_al2012}).
+for HERA.  The full-Stokes sky maps resulting from this pipeline
+are used to directly subtract foregrounds
+and as data products themselves.  UPenn leads
+the characterization of the polarized sky and the
+distribution of rotation measures.  Data from HERA, PAPER, and the MWA are used
+to update source catalogs and a Global
+Sky Model \Mycitep{deoliveira2008}. %providing reference sky models for both
+%HERA and other low-frequency instruments. %The goal will be to generate a model
+%suitable for direct subtraction of foregrounds.
+
+% XXX need more data products here
+
+%\compress
+%\subsubsection{Power spectrum and related measurements}
+
+{\bf Power Spectra}. UC Berkeley leads early power spectrum measurements using the conservative
+delay-spectrum approach used in PAPER.  %This
+%visibility-based approach, when coupled with a foreground-avoidance strategy,
+%is sufficient for an unambiguous early detection of the 21~cm
+%power spectrum should results from current-generation experiments prove
+%marginal (Fig. \ref{fig:eor_pspec} and Table \ref{tab:signif}).  
+Development of
+the power-spectrum pipeline will focus on incorporating
+a fully covariant description of the foreground wedge
+\Mycitep{liu_tegmark2011,dillon_et_al2013a}, as well as
+exploring Bayesian techniques for power spectrum estimation, building on the considerable
+progress already made on Gibbs-sampling imaging for interferometers \Mycitep{sutter_et_al2014}.
+In addition to the statistical foreground mitigation techniques provided by optimal quadratic
+power-spectrum estimation, U. Washington leads the application of 
+foreground modeling for expanding the range of modes available for power spectrum analysis.
+Techniques developed should be equally applicable
+to measurements in both the reionization epoch and at higher redshifts.  MIT leads the application
+of these tools for pre-reionization science.
+
+{\bf Simulation}. To provide verification of power spectrum constraints, ASU leads the
+development of an end-to-end simulator of the full HERA instrument, incorporating
+foregrounds and reionization models to output visibilities.  UCLA also leads
+a parametrized reionization simulation effort that, coupled with the instrument
+simulator, is critical for connecting power-spectrum constraints to the estimation
+of the underlying astrophysical models.
+
+%\compress
+%\subsubsection{HI Imaging}
+
+{\bf HI Imaging}. In addition to supplying models for suppressing foregrounds,
+deep imaging aims to map HI emission during reionization (see \S\ref{sec:science}).
+Initially, conservative filtering of the foreground wedge is applied to data that
+are imaged.  These results are compared with the residuals from imaging-based
+foreground suppression to explore the limitations of 
+precision foreground imaging and removal.  As imaging-based techniques mature, the resultant
+maps are released for cross-correlation with other probes of reionization.
+
+%\subsubsection{Aspirations [move to facilities section?]}
+%
+%i. FFT correlator
+%ii. other
+
+%\compress
+%\subsection{Schedule} % 0.5 page
+% De Boer
+
+%\begin{table}[t]
+%\centering
+%\caption{Schedule Overview}
+%\label{tab:scheduleSummary}
+%\begin{tabular}{| p{1.5in} | p{1.5in} | p{1.5in} | p{1.5in} |}\hline
+%\textbf{Year 1}:  FY2015   &  \textbf{Year 2}:  FY2016  &  \textbf{Year 3}:  FY2017 & \textbf{Year 4}:  FY2018 \\ \hline
+%\raggedright{Infrastructure, prototyping and contract preparation.} &
+%\raggedright{Hardware commissioning and deep foreground survey.} &
+%\raggedright{HERA-127 observations:  detecting the rise and fall of reionization.} &
+%HERA-331 observations:  measuring the evolution of the first galaxies. \\ \hline  %It won't let me make this one raggedright!
+%\begin{itemize}[noitemsep,nolistsep,leftmargin=12pt]
+%\item Basic infrastructure
+%\item GB, SA prototypes
+%\item Design package
+%\item Polarization software
+%\end{itemize} &
+%\begin{itemize}[noitemsep,nolistsep,leftmargin=12pt]
+%\item HERA beam pattern
+%\item HERA-127
+%\item Delay-spectrum/FHD software
+%\item Analysis pipelines
+%\end{itemize} &
+%\begin{itemize}[noitemsep,nolistsep,leftmargin=12pt]
+%\item HERA-127 observations
+%\item Electronics upgrade
+%\item HERA-331
+%\item HERA-127 analysis
+%\end{itemize} &
+%\begin{itemize}[noitemsep,nolistsep,leftmargin=12pt]
+%\item HERA-331 observations
+%\item Cross-correlations
+%\item Cosmological simulations
+%\end{itemize} \\ \hline
+%\end{tabular}
+%\end{table}
+
+% need to discuss in a few sentences the
+% relation to other experiments in the same time frame.
+
+%\begin{itemize}[noitemsep,nolistsep]
+%\item{Year 1:} Infrastructure, Prototyping, Contract Preparation (FY 2015).      
+%\begin{itemize}[noitemsep,nolistsep]
+%\item Install basic infrastructure (ground leveling, power, network connectivity) for new infrastructure
+%\item Incorporate existing PAPER-128 antennas, correlator, and housing container.
+%\item Install additional prototypes in Green Bank and test.
+%\item Define final design package for element construction
+%\item Start developing improved HERA baluns, receivers, feeds, nodes, and in-situ antenna calibration system.
+%\item Continue delay-spectrum, FHD, and optimal estimator software development.
+%\item Develop polarization capable software for beam/leakage studies.
+%\end{itemize}
+%\item{Year 2:} Hardware Commissioning and Deep Foreground Survey (FY 2016).  
+%\begin{itemize}[noitemsep,nolistsep]
+%\item Observations using PAPER antennas in an imaging configuration.
+%\item Determine on-sky beam response of HERA antennas to facilitate future source subtraction efforts.
+%\item Finalize site infrastructure (high-bandwidth optical network, surveying, trenching).
+%\item Commission new feeds, receivers, nodes, and calibration systems in Green Bank and SA.
+%\item Complete HERA-127 construction
+%\item Initial delay-spectrum, FHD, and optimal estimator software ready for HERA 127 analysis.
+%\item Full end-to-end simulations of analysis pipelines.
+%\end{itemize}
+%\item{Year 3:} HERA 127 and Detecting the Rise and Fall of Reionization (FY 2017).  
+%\begin{itemize}[noitemsep,nolistsep]
+%\item HERA 127 complete. Science observations begin using the PAPER correlator.
+%\item Begin deployment of HERA 331. Install new node electronics and a 331-element, GPU-based correlator in the Karoo Array Processing Building (KAPB).
+%\item Install new data storage infrastructure in the KAPB.
+%\item Upgrade the UPenn analysis cluster.
+%\item Finish construction of HERA 331
+%\item Apply proven delay-spectrum analysis techniques to HERA 127 observations to constrain the timing and duration of reionization.
+%\item Multi-pipeline analysis of HERA 127 observations to compare pipeline strengths/weaknesses.
+%\end{itemize}
+%\item{Year 4:} HERA 331 and Measuring the Evolution of the First Galaxies (FY 2018).           
+%\begin{itemize}[noitemsep,nolistsep]
+%\item HERA 331 complete.
+Begin science observations Oct.
+2017.\item Begin multi-pipeline analysis of data to characterize the evolution of the power spectrum and determining properties of the first galaxies.
+%\item Continue analysis software development: optimize existing algorithms and develop imaging-based subtraction techniques for expanding the EoR window
+%\item Create high-fidelity maps/data cubes of the HERA field for cross-correlation studies
+%\item Cosmological simulations for extracting reionization physics from power spectrum measurements
+%\end{itemize}
+%\end{itemize}
 
@@
 
+\begin{figure}[t]
+\vspace*{2mm}
+\begin{center}
+\includegraphics[width=\columnwidth]{plots/costing.png}
+\end{center}
+\caption{Costing  \label{fig:costing}}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+        \includegraphics[height=1.5in]{plots/new_antenna_closeup.jpg}
+        \includegraphics[height=1.5in]{plots/Engineering/recv_node.png}
+        \includegraphics[height=1.5in]{plots/Engineering/digital.png}
+    \Caption{-0.1in}{0.99}{-0.1in}{\small
+    Existing components re-used in the HERA design include:
+    the PAPER dipole antenna (left), 
+    receivers in a node module (center), and
+    the 128-element correlator deployed in the Karoo (right).
+}\label{fig:components}
+\end{figure}
+
 %% TWO-COLUMN FIGURES
+
+\begin{figure}[t]
+\vspace*{2mm}
+\begin{center}
+\includegraphics[width=12cm]{plots/HERA_high_level_block_diagram.png}
+\end{center}
+\caption{System  \label{fig:systemOverview}}
+\end{figure}
+
+\begin{figure}[t]
+\vspace*{2mm}
+\begin{center}
+\includegraphics[width=12cm]{plots/config.png}
+\end{center}
+\caption{Configuration  \label{fig:config}}
+\end{figure}
 
@@
 %\end{table}
+\begin{table}[t]
+\small
+ \centering
+ \begin{tabular}{c||r||r|r} 
+%\begin{deluxetable}{c||r||r|r}
+%\tabletypesize{\small}
+%\tablecaption{\small
+\hline
+%\startdata
+Instrument & Collecting Area (m$^2$) & Foreground avoidance & Foreground modeling \\
+\hline
+PAPER & 528 & 1.93 & 8.86 \\
+MWA & 896 & 2.46 & 6.40 \\
+LOFAR NL Core & 35,762 & 2.76 & 17.37 \\
+\textbf{HERA-127} & \textbf{19,500} & \textbf{10.88} & \textbf{35.65} \\
+\textbf{HERA-331} & \textbf{50,900} & \textbf{25.44} & \textbf{87.20} \\
+SKA1 Low Core & 833,190 & 97.92 & 284.85 \\
+%\enddata
+\end{tabular}
+\Caption{-0.1in}{0.99}{-0.1in}{Power spectrum signal-to-noise (``number of sigmas") at $z=9.5$ for various instruments, adapted from \Mycitet{pober_et_al2014}.
+By leveraging a filled, redundant configuration of dishes with high collecting area, HERA-331 allows high-significance power spectrum measurements using current foreground avoidance techniques, with further enhancements possible with likely advances in foreground modeling.}
+\label{tab:signif}
+\end{table}
+
+\begin{table}[t]
+\small
+\begin{center}
+% \begin{deluxetable}{lcc}
+% \tabletypesize{\small}
+% \tablecaption{HERA-331 basic parameters.}
+% \tablehead{\colhead{Parameter} & \colhead{Design} & \colhead{Performance at 150 MHz}}
+%    \startdata
+\begin{tabular}{l | c | c}
+\hline
+Parameter & Design & Performance\\
+\hline
+    Element diameter / FoV & 14 m & 9\arcdeg \\ 
+    %Total collecting area & 54186 m$^2$ \\
+    Min baseline length / largest scale & 14.6 m & 7.8\arcdeg \\
+    Max core baseline length / synthesized beam & 306.6 m & 24\arcmin \\ 
+    Max outrigger baseline length  & 1066.5 m & 9\arcmin \\
+    Frequency / redshift range  & 50 - 250 MHz digitized \\
+    & 70 - 230 MHz useable & 19.2 - 5.2 \\ 
+    & 100 MHz correlated & \\
+    Spectral channel width & 97.7 kHz & \\    
+    System temperature / sensitivity & $100 + 120 (\nu/\rm{150~MHz})^{-2.55}$ K 
+    & 50 $\mu \rm{Jy}~\rm{beam}^{-1}~\sqrt{\rm{hour}}$ \\
+    \hline
+%     At 150 MHz ($z=8.5$): & \\
+%     ~{   }Naturally weighted synthesized beam FWHM & $24\arcmin$ \\
+%     ~{   }Uniformly weighted synthesized beam FWHM & $9\arcmin$ \\
+%     ~{   }Field of view FWHM & 9\arcdeg \\
+%     ~{   }Point source RMS & 50 $\mu$Jy in 100 hrs \\
+%  \enddata
+%    \hline
+ %\end{deluxetable}
+\end{tabular}
+\Caption{-0.1in}{0.99}{-0.4in}{HERA-331 basic parameters.  Design parameters are connected to the derived instrument performance at 150 MHz.}
+\label{tab:BasicParameters}
+\vspace{.2in}
+%\end{deluxetable}
+\end{center}
+\end{table}
+
+%\end{deluxetable}
 

### 2014-07-02 hera-concept2013.tex
starting to clean up e.g. cite etc

@@
 \def\xHI{{x_{\rm\HI}}}
+\def\arcdeg{{$^o$}}
+\def\arcmin{{$''$}}
 \begin{document}
@@
 %% These dates will be inserted by the Publication Production Office during the typesetting process.
+
+% Intro
+% Background
+% Scientific Motivation
+% Specifications
+% System Overview
+% Data Processing
+% Site?
+% Conclusion
 
@@
 wedge must be projected out of our measurements.
-As summarized in Table \ref{tab:signif} (see \Mycitealt{pober_et_al2014} for more details),
+As summarized in Table \ref{tab:signif} (see \citealt{pober_et_al2014} for more details),
 based on these assumptions, current instruments are likely to achieve,
@@
 %%		\includegraphics[width=\textwidth]{plots/paper_element.jpg}
-%%		\Caption{-0.2in}{0.99}{-0.2in}{The PAPER element (provides a clean instrumental response as a function
+%%		\caption{-0.2in}{0.99}{-0.2in}{The PAPER element (provides a clean instrumental response as a function
 %%		of frequency \citep{parsons_et_al2010,parsons_et_al2012b}, which is crucial to
@@
 %%		\includegraphics[height=1.75in]{plots/hera_dish.png}
-%%		\Caption{-0.2in}{0.99}{-0.2in}{A 14m dish designed around the feed dramatically improves sensitivity while
+%%		\caption{-0.2in}{0.99}{-0.2in}{A 14m dish designed around the feed dramatically improves sensitivity while
 %%		constraining the path length and amplitude of reflections to ensure that foreground 
@@
 %%	\end{subfigure}
-%%\Caption{-0.2in}{0.99}{-0.2in}{PAPER and HERA elements}
+%%\caption{-0.2in}{0.99}{-0.2in}{PAPER and HERA elements}
 %%\label{fig:hera_dish}
@@
 antenna separations of $8\lambda \approx 15$m are
-well-behaved to current limits \Mycitep{pober_et_al2013,parsons_et_al2013}. This influences our
+well-behaved to current limits \citep{pober_et_al2013,parsons_et_al2013}. This influences our
 choice of a 14-m dish diameter,
@@
 \S\ref{sec:Lessons}, redundant configurations have been employed by PAPER to
-boost sensitivity \Mycitep{parsons_et_al2012a}, with the added benefit that they
+boost sensitivity \citep{parsons_et_al2012a}, with the added benefit that they
 greatly facilitate fast and accurate calibration
-\Mycitep{liu_et_al2010,parsons_et_al2013}.  Placing HERA elements in a
+\citep{liu_et_al2010,parsons_et_al2013}.  Placing HERA elements in a
 grid allows certain construction components to be re-used between antennas,
@@
 (Fig. \ref{fig:components}, left and center) that are currently being
-used to good effect in PAPER-128 \Mycitep{parsons_et_al2010}.  Existing PAPER feeds are attached to new
+used to good effect in PAPER-128 \citep{parsons_et_al2010}.  Existing PAPER feeds are attached to new
 reflector screens that are suspended over the elements, but otherwise remain unmodified.
@@
 \includegraphics[width=6.5in]{plots/casper_boards.jpg}
-\Caption{-0.3in}{0.99}{-0.15in}{\small
+\caption{-0.3in}{0.99}{-0.15in}{\small
 Five generations of CASPER technology (progressing left to right) have been used to rapidly
@@
 easily upgraded for HERA, incorporating new technology
-\Mycitep{parsons_et_al2006,parsons_et_al2008}.
+\citep{parsons_et_al2006,parsons_et_al2008}.
 }\label{fig:casper_boards}
@@
 expensive, and risky aspects of developing a radio interferometer, this is no longer the case.
-CASPER \Mycitep{parsons_et_al2006}
+CASPER \citep{parsons_et_al2006}
 open-sourced the development of digital signal processing engines for astronomy and
@@
 HERA efforts continue this incremental development cycle, following a packet-switched
-correlator architecture \Mycitep{parsons_et_al2008} that has been
+correlator architecture \citep{parsons_et_al2008} that has been
 extended in recent PAPER and LEDA deployments (Fig. \ref{fig:components}, right)
@@
 and deployment of three data storage systems:
-\begin{itemize}[noitemsep,nolistsep]
+\begin{itemize}%[noitemsep,nolistsep]
 \item a 1.5 PB system deployed in the KAPB for archiving all 1.2 PB of raw data and 60 TB of compressed data, 
@@
 The data compression scheme at the heart of the data management system
-has been implemented for PAPER (\Mycitealt{parsons_et_al2013},
+has been implemented for PAPER (\citealt{parsons_et_al2013},
 Appendix A), and is applied to HERA visibilities to reduce data volume by
@@
 capabilities.  This compression technique, which is based on delay/delay-rate
-filtering \Mycitep{parsons_backer2009}, is applied uniformly to all visibilities
+filtering \citep{parsons_backer2009}, is applied uniformly to all visibilities
 in the array, does not require (or produce) detailed calibration
@@
 The U. of Washington team leads the development of HERA's Monitor and Control (M\&C) system,
-which is a straightforward port of a similar system used on the MWA \Mycitep{tingay_et_al2013}.
+which is a straightforward port of a similar system used on the MWA \citep{tingay_et_al2013}.
 This system is
@@
 of a high dynamic-range imaging pipeline
-based on the Fast Holographic Deconvolution (FHD; \Mycitealt{sullivan_et_al2012}).
+based on the Fast Holographic Deconvolution (FHD; \citealt{sullivan_et_al2012}).
 for HERA.  The full-Stokes sky maps resulting from this pipeline
@@
 to update source catalogs and a Global
-Sky Model \Mycitep{deoliveira2008}. %providing reference sky models for both
+Sky Model \citep{deoliveira2008}. %providing reference sky models for both
 %HERA and other low-frequency instruments. %The goal will be to generate a model
@@
 a fully covariant description of the foreground wedge
-\Mycitep{liu_tegmark2011,dillon_et_al2013a}, as well as
+\citep{liu_tegmark2011,dillon_et_al2013a}, as well as
 exploring Bayesian techniques for power spectrum estimation, building on the considerable
-progress already made on Gibbs-sampling imaging for interferometers \Mycitep{sutter_et_al2014}.
+progress already made on Gibbs-sampling imaging for interferometers \citep{sutter_et_al2014}.
 In addition to the statistical foreground mitigation techniques provided by optimal quadratic
@@
 
-\begin{thebibliography}{}
-
-\bibitem[AUTHOR(YEAR)]{LABEL}
-REFERENCE 1
-
-\bibitem[AUTHOR(YEAR)]{LABEL}
-REFERENCE 2
-
-...
-
-\end{thebibliography}
+\bibliographystyle{copernicus}
+\bibliography(biblio)
+
+%\begin{thebibliography}{}
+%
+%\bibitem[AUTHOR(YEAR)]{LABEL}
+%REFERENCE 1
+%
+%\bibitem[AUTHOR(YEAR)]{LABEL}
+%REFERENCE 2
+%
+%...
+%
+%\end{thebibliography}
 
@@
         \includegraphics[height=1.5in]{plots/Engineering/digital.png}
-    \Caption{-0.1in}{0.99}{-0.1in}{\small
+    \caption{-0.1in}{0.99}{-0.1in}{\small
     Existing components re-used in the HERA design include:
@@
 %\end{table}
-\begin{table}[t]
+\begin{table*}[t]
 \small
@@
 \end{tabular}
-\Caption{-0.1in}{0.99}{-0.1in}{Power spectrum signal-to-noise (``number of sigmas") at $z=9.5$ for various instruments, adapted from \Mycitet{pober_et_al2014}.
+\caption{Power spectrum signal-to-noise (``number of sigmas") at $z=9.5$ for various instruments, adapted from \citet{pober_et_al2014}.
 By leveraging a filled, redundant configuration of dishes with high collecting area, HERA-331 allows high-significance power spectrum measurements using current foreground avoidance techniques, with further enhancements possible with likely advances in foreground modeling.}
 \label{tab:signif}
-\end{table}
-
-\begin{table}[t]
+\end{table*}
+
+Print this is some text in between.
+
+\begin{table*}[t]
 \small
-\begin{center}
+%\begin{center}
 % \begin{deluxetable}{lcc}
@@
 %    \startdata
-\begin{tabular}{l | c | c}
-\hline
-Parameter & Design & Performance\\
-\hline
+\begin{tabular}{l | c | c} \hline
+Parameter & Design & Performance\\  \hline
     Element diameter / FoV & 14 m & 9\arcdeg \\ 
@@
 \end{tabular}
-\Caption{-0.1in}{0.99}{-0.4in}{HERA-331 basic parameters.  Design parameters are connected to the derived instrument performance at 150 MHz.}
+\caption{HERA-331 basic parameters.  Design parameters are connected to the derived instrument performance at 150 MHz.}
 \label{tab:BasicParameters}
@@
 %\end{deluxetable}
-\end{center}
-\end{table}
+%\end{center}
+\end{table*}
 

### 2015-08-11 eorsens.py
commit of everything before I start hacking for new version

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+sens = """PAPER 128  &  1.56  &  -  &  4.46  &  - 
+MWA 128      &  0.66  & 0.86  &  2.50  &  3.15 
+LOFAR (core) & 0.70  & 1.90  & 7.48  &  12.22 
+HERA 37         & 5.67  &  -  & 15.46   &  - 
+HERA 331       & 38.75 & -  &  111.69  &  - 
+MWA 256         &2.40   &  2.81  &  8.28  & 9.64 
+SKA1 Low      &21.23   & 26.92  & 139.07  & 115.13"""
+
+sens = sens.split('\n')
+arrnames = []
+arrsens = []
+for line in sens:
+    data=line.split('&')
+    arrnames.append(data[0].strip())
+    del(data[0])
+    varr = []
+    for v in data:
+        if v.strip()=='-':
+            varr.append(0.0)
+        else:
+            varr.append(float(v))
+    arrsens.append(varr)
+arrsens = np.array(arrsens)
+print arrnames
+print arrsens
+
+gridlines = [0.6,0.7,0.8,0.9,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,20.0,30.0,40.0,50.0,60.0,70.0,80.0,90.0,100.0]
+gridlines = [1.0,10.0,100.0]
+for g in gridlines:
+    plt.plot([-1.5,13.5],[g,g],'k:')
+
+w = 0.4
+c = ['b','c','m','r']
+labels = ['Avoid(drift)','Subtract(drift)','Avoid(track)','Subtract(track)']
+cols = [0,2,1,3]
+ll = np.arange(len(arrsens[:,0]))*2-0.5+w/2.0
+for i in range(len(cols)):
+    plt.bar(ll+i*w*1.05,arrsens[:,cols[i]],color=c[i],width=w,label=labels[i])
+plt.yscale('log')
+plt.axis([-1.5,13.5,0.5,140.0])
+plt.legend(loc='upper left')
+xt = [0,2,4,6,8,10,12]
+plt.xticks(xt,arrnames,rotation=70)
+plt.plot([5.55,5.55],[0.5,140],'k',linewidth=3)
+
+

### 2015-08-12 manuscript.tex
renamed tex files to new versions --> active one is manuscript.tex

% Template file for ICEAA13
%
\RequirePackage{ifpdf}
\documentclass{ICEAA-IEEE_APWC}
\usepackage{multirow}

\newcommand{\kvec}{{\bf k}}
\newcommand{\bvec}{{\bf b}}
\newcommand{\shat}{{\hat s}}
\newcommand{\kpr}{{k_\perp}}
\newcommand{\kvpr}{{\kvec_\perp}}
\newcommand{\kpar}{{k_\parallel}}
\newcommand{\AI}{{\langle\tilde A*\tilde I\rangle}}
\newcommand{\AItau}{{\AI(\tau)}}
\newcommand{\hMpci}{{h~{\rm Mpc}^{-1}}}
\newcommand{\inch}{$^{\prime\prime}$}
\newcommand{\foot}{$^{\prime}$}
\renewcommand{\deg}{^\circ}

% In order to produce the pdf document for paper submission either:
% a) process with pdfLaTeX, or
% b) process with LaTeX, dvips, ps2pdf
% In both cases comment out the document class option as indicated above

\newcommand\aj{AJ}%
          % Astronomical Journal
\newcommand\actaa{\ref@jnl{Acta Astron.}}%
  % Acta Astronomica
\newcommand\araa{\ref@jnl{ARA\&A}}%
          % Annual Review of Astron and Astrophys
\newcommand\apj{ApJ}%
          % Astrophysical Journal
\newcommand\apjl{\ref@jnl{ApJ}}%
          % Astrophysical Journal, Letters
\newcommand\apjs{\ref@jnl{ApJS}}%
          % Astrophysical Journal, Supplement
\newcommand\ao{\ref@jnl{Appl.~Opt.}}%
          % Applied Optics
\newcommand\apss{\ref@jnl{Ap\&SS}}%
          % Astrophysics and Space Science
\newcommand\aap{\ref@jnl{A\&A}}%
          % Astronomy and Astrophysics
\newcommand\aapr{\ref@jnl{A\&A~Rev.}}%
          % Astronomy and Astrophysics Reviews
\newcommand\aaps{\ref@jnl{A\&AS}}%
          % Astronomy and Astrophysics, Supplement
\newcommand\icarus{\ref@jnl{Icarus}}%
  % Icarus
\newcommand\mnras{\ref@jnl{MNRAS}}%
          % Monthly Notices of the RAS
\newcommand\prc{\ref@jnl{Phys.~Rev.~C}}%
          % Physical Review C
\newcommand\prd{\ref@jnl{Phys.~Rev.~D}}%
          % Physical Review D
\newcommand\pre{\ref@jnl{Phys.~Rev.~E}}%
          % Physical Review E
\newcommand\prl{\ref@jnl{Phys.~Rev.~Lett.}}%
          % Physical Review Letters
\newcommand\pasa{\ref@jnl{PASA}}%
  % Publications of the Astron. Soc. of Australia
\newcommand\pasp{\ref@jnl{PASP}}%
          % Publications of the ASP

\title{The Hydrogen Epoch of Reionization Array (HERA)}

% Use \thanks{} for each affiliation.
% If the authors have the same affiliation, use \thanks[n] to
% refer to the same affiliation of the n:th \thanks command
\author{D. R. DeBoer\thanks{Radio Astronomy Laboratory, University of California, Berkeley, CA 94720, USA, e-mail: ddeboer@berkeley.edu}, 
on behalf of the HERA collaboration
}
\begin{document}
\maketitle

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR). During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.  This paper will provide an overview of the project and describe the design of the HERA receiving element.

\end{abstract}

\section{Introduction}
\label{sec:intro}
We have garnered a deep understanding of the large scale structure of the very early Universe by observing the faint afterglow of the cosmic microwave background (CMB).  The CMB signal arises due to the recombination of hydrogen gas which allows the radiation to stream forth.  It represents one slice in time when the faint imprint of the nearly homogeneous inflationary Universe may be seen.

We also have a good understanding of the complicated structure and denizens of our current Universe.  In between however, we have very little information of how our Universe evolved to its current structure.  It is as if we have one detailed snapshot from one day in the first days of our baby's life, and then a Facebook page of our child as a young adult with very little in between except faint rumors and a headline or two.

This intervening period has essentially four phases which may be called the Dark Ages, the Cosmic Dawn, the Epoch of Reionization (EOR) and the Rise of the Galaxies.  The time of the Cosmic Dawn saw the first large-scale gravitational collapse of structures, which eventually heated up the neutral intergalactic hydrogen gas until it again ionized.  This phase change is called the Epoch of Reionization and it represents the last phase change of the Universe as a body.  We may observe this epoch over its extent by observing the red-shifted hydrogen gas.  In our analogy, this is akin to following our child's activities and growth through a good part of their formative years.

Detecting, characterizing and ultimately imaging the Epoch of Reionization is a key goal for the astronomy and astrophysics community. There are many references on the theoretical, instrumental and observational foundations, see for example one of the latest instrument papers and references therein ({\em e.g.} {\cite{2015arXiv150206016A}}).  Current projects (PAPER (http://eor.berkeley.edu), MWA (http://mwatelescope.org), LOFAR (http://www.lofar.org)) are striving to make the first detection of the statistical power spectrum of the signal, however current best limits still fall above even optimistic predictions of its intrinsic strength.  While these projects are still taking data, it is recognized that an optimized array based on our new understanding of the signal characteristics is needed to make a strong detection and begin to characterize this signal over multiple scales and redshifts.  These results would also inform design elements for the Square Kilometer Array (http://www.skatelescope.org).

The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the EOR, roughly 0.3 - 1 billion years after the Big Bang. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.

\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{hera331_Bpers.png} }
\caption{\small Perspective cartoon of the 331-element core within a 305-meter circle.
\label{fig:config}}
\end{figure}

HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection.  The key is to produce inexpensive collecting area optimized for the spatial scales needed to detect the EOR.  This has resulted in a close-packed array of transit 14-meter dishes (Fig. \ref{fig:config}), with the first elements currently under construction at the South African Karoo Astronomy Reserve, the current location of the Donald C. Backer Precision Array to Probe the Epoch of Reionization (PAPER) array.

Beginning with the current NSF funding, HERA will rollout in a staged deployment.  The first phase extending to August 2016 will deploy 37 elements that will test the HERA element design for performance and manufacturability at location.  These 37 elements will more than double the sensitivity of the 128-element PAPER array.  The next phase extending into 2018 will build out to 127 elements, which should provide a very robust detection of the EOR signal.  Finally, extending into 2020, HERA will build out to 352 elements to further EOR science as a function of redshift and spatial scale, potentially producing the first images of the EOR.

HERA is a collaboration of the following partner institutions:  Arizona State University, University of California Berkeley, University of California Los Angeles, University of Cambridge, University of KwaZulu Natal, Massachusetts Institute of Technology, National Radio Astronomy Observatory, University of Pennsylvania, SKA-South Africa, and
University of Washington.

\section{Measuring the EOR}
\label{sec:eormeas}
Due to the expansion of the Universe, we can identify and measure the early Universe via the red-shift of spectral lines.  The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.  Frequency therefore equates to a volume along the site of the telescope as well as in dating when the event occurred.

Initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.  The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.  Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of its properties to achieve this more difficult goal.

Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.  We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment (see {\em e.g.} \cite{2012ApJ...756..165P}):
\begin{equation}
{\hat P}(\kvec) \approx \left(\frac{X^2Y}{4k_B^2}\right) \frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} 
\end{equation}
where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated squared beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.  The terms not in parentheses are instrumental terms, as opposed to constants and cosmological parameters.

Though the desired measurement is over a 3-D volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kpr$ is determined by the antenna baseline and $\kpar$ by the frequency.  This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything not the EOR) contaminate the signal of interest from where they don't.  This contaminated phase space may be shown to be a wedge-shaped region in $\kpr$:$\kpar$ space:
\begin{equation}
|\kpar| \le \frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
\end{equation}
where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response.  
A key issue to measuring the EOR power spectrum is to understand and minimize such effects.  For an analytical derivation, see \cite{2012ApJ...756..165P,vedantham_et_al2012,liu_et_al2014b},  in simulation see \cite{datta_et_al2010,hazelton_et_al2013} and for observations
\cite{2013ApJ...768L..36P,2015arXiv150601026P,2014ApJ...788..106P,2015arXiv150206016A}.

\section{HERA System}
\label{sec:system}
HERA will operate between 70 MHz and 220 MHz, however the first phases (up to 127 elements) will operate within the current PAPER bandwidth of 110-190 MHz.  The full band is directly sampled and correlated, making for a very simple signal path requiring only gain.  Initially, the exact PAPER signal chain will be used, which brings the RF back to a central container on coaxial cable where it is digitized and processed.  A new modular digitizer (Smart Network ADC Processor, or SNAP) has been developed that will be deployed in the field and will shorten the amount of RF cable.  The digitized signal will be conveyed back to a central building for processing via 10 Gbit ethernet over fiber-optic cable.   The SNAP board is being developed as part of the CASPER suite of hardware components (https://casper.berkeley.edu/wiki/DAB-HERALD).  A simple block diagram of the two-phase system is shown in Fig. \ref{fig:system}.

\begin{figure}[t]
\centerline{
\includegraphics[width=5.5cm]{system.png} }
\caption{\small High-level block diagram of the HERA system.  The dashed lines indicate the first phase while solid are the final design.  Black lines are coaxial cable and orange are ethernet over fiber.
\label{fig:system}}
\end{figure}

The configuration is optimized for power spectrum detections, which favors as large of a fill-factor as possible.  The configuration is therefore a packed hexagonal array with a 14.6-meter pitch with a 14-meter antenna (see Fig \ref{fig:config}).  Additionally, there will be 21 outriggers out to 1.2km to improve the resolution

%\begin{table}[htpb]
%\centerline{
%\begin{tabular}{|l||c|c||c|c||}\hline
% \multirow{2}{*}{Instrument}                 & \multicolumn{2}{|c|}{Avoidance} & \multicolumn{2}{|c|}{Subtraction} \\ \cline{2-5}
%                      & Drift  & Track  & Drift  & Track \\ \hline
%PAPER 128  &  1.56  &  -  &  4.46  &  - \\ \hline
%MWA 128      &  0.66  & 0.86  &  2.50  &  3.15 \\ \hline
%LOFAR (core) & 0.70  & 1.90  & 7.48  &  12.22 \\ \hline \hline
%HERA 37         & 5.67  &  -  & 15.46   &  - \\ \hline
%HERA 331       & 38.75 & -  &  111.69  &  - \\ \hline
%MWA 256         &2.40   &  2.81  &  8.28  & 9.64 \\ \hline
%SKA1 Low      &21.23   & 26.92  & 139.07  & 115.13 \\ \hline
%\end{tabular}}
%\caption{\small Parameters and their values; units are spaced from the
%corresponding measure by an unbreakable thin space and should
%always be in upright font.\label{tab:sensitivity}}
%\end{table}

\section{HERA Element}
\label{sec:element}
The design of the HERA element is discussed in \cite{heraMemo5}.  The design principles are three-fold:
\begin{itemize}
\item optimize for the delay-spectrum technique of measuring the EOR power spectrum,
\item minimize costs, and
\item the experiment has a limited lifetime of about five years.
\end{itemize}
The first item primarily means that chromatic effects corresponding to delays appropriate for the measurement described above must be below the expected signal level, which essentially determines the focal length.  The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.  And the third items constrains the construction materials and methods and the operational model.

The approach used first sets a specification on the magnitude of the reflections within the dish on the delays of interest.  Reference \cite{heraMemo5} derives this to be that the reflections must be down by 60 dB on timescales greater then 60 nanoseconds.  
Next, using a feed pattern based on the existing PAPER dipole antenna and a fairly complete analytical model of a prime focus antenna, the effective area was found to be maximized for $f/D \approx 0.32$.  Figure \ref{fig:costfig} shows the vertex/feed round-trip travel times for one, two and three reflections with the 60 ns timescale indicated by the black dashed line.

To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity for this measurement, which is shown in Figure \ref{fig:costfig}.  By assuming canonical values of 331 14-meter antennas, the number of elements needed for a given diameter to match the sensitivity is
\begin{equation}
N_a = 26918D_a^{-\frac{5}{3}}
\end{equation}
Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve, which becomes quite flat for diameters greater than about 14m, which also happens to be the diameter where two reflections between the feed and vertex coincide with the 60 ns specification.  An addition high-level analysis of reflection coefficients within the structure also indicate that a diameter of less than 15 meters is preferred.  The diameter of 14 meters was therefore chosen for the element.

\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{costfig.png} 
}
\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).  The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
\label{fig:costfig}}
\end{figure}

A full-sized prototype of this system was built and measured using a network analyzer in time-domain mode, which meets the ``60 by 60 '' specification (Fig \ref{fig:refl}). (\cite{heraMemo5}).

\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{reflections.png} 
}
\caption{\small Network analyzer measurements of reflected power versus delay time of the full-size HERA prototype.
\label{fig:refl}}
\end{figure}

\section{Conclusion}
A system design meeting the above design principles was presented, which comprises 331 14-meter elements in a compact hexagonal grid and 21 outriggers out to a 1.2km maximum baseline.  HERA uses the demonstrated foreground avoidance technique and should provide a very robust detection of the EOR signal.  Additional work will continue on HERA and other arrays to get the first detection as well as to perfect techniques allowing us to get to the other spatial scales and image the EOR.  Figure \ref{fig:eorsense} compares the sensitivity of existing arrays (left of the thick vertical line) using the different techniques, along with future HERA and SKA arrays.  HERA has the strong potential to get a robust detection in the near-term and characterize the EOR and signal processing to feed into the SKA to get a more complete picture of this period in the life of our early Universe.

\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{eorsens.png} 
}
\caption{\small Sensitivity calculations for present and future arrays  (separated by the vertical thick black line) investigating the EOR.
The arrays are mentioned in Sec. \ref{sec:intro}.  The four bars per array refer to different observing and analysis techniques.  The "avoidance" 
methods are largely demonstrated while the "subtraction" ones are still under development.  Note that PAPER/HERA do not track.
\label{fig:eorsense}}
\end{figure}
\section*{Acknowledgments}
This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. 1440343.
\bibliographystyle{plain}
\bibliography{mybibdesk}

\end{document}

### 2015-08-12 orig_hera_concept.tex
renamed tex files to new versions --> active one is manuscript.tex

--- 2014-07-02 hera-concept2013.tex
@@
 %% ---------------------------------
-%% This template should be used for the following class files: copernicus.cls, copernicus2.cls, copernicus_discussions.cls
-%% The class files, the Copernicus LaTeX Manual with detailed explanations regarding the comments
-%% and some style files are bundled in the Copernicus Latex Package which can be downloaded from the different journal webpages.
-%% For further assistance please contact the Publication Production Office (production@copernicus.org).
+%% This template should be used for copernicus.cls
+%% The class file and some style files are bundled in the Copernicus Latex Package which can be downloaded from the different journal webpages.
+%% For further assistance please contact the Copernicus Publications at: publications@copernicus.org
 %% http://publications.copernicus.org
 
-%% Differing commands regarding the specific class files are highlighted.
-
-%% copernicus.cls
-\documentclass[ars]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus}
-
-%% copernicus2.cls
-%\documentclass[journal abbreviation]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus2}
-
-%% copernicus_discussions.cls
-%\documentclass[ars, hvmath, online]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus_discussions}
-\usepackage{amsmath}
+\documentclass[ars]{copernicus}
+
+%\usepackage{amsmath}
 \def\kperp{k_{\bot}}
@@
 \newcommand{\simgt}{\stackrel{>}{_{\sim}}}
+\newcommand{\mnras}{Month. Not. Royal Astron. Soc.}
+\newcommand{\apj}{Astrophys. J.}
+\newcommand{\apjl}{Astrophys. J. Lett.}
+\newcommand{\pasa}{Proc. Astron. Soc. Aust.}
+\newcommand{\pasp}{Proc. Astron. Soc. Pac.}
+\newcommand{\aj}{Astron. J.}
+\newcommand{\prd}{Unknown}
+\def\eg{{\em e.g.~}}
 \def\kperp{k_{\bot}}
@@
 \def\arcmin{{$''$}}
+
 \begin{document}
@@
 
-\title{The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA)}
+\title{The Hydrogen Epoch of Reionization Array (HERA)}
 
@@
 \begin{abstract}
-The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged
+Hydrogen Epoch of Reionization Array  is a staged
 experiment that uses the unique properties of the 21-cm line from neutral
@@
 and black holes, their influence on the intergalactic medium (IGM), and
-cosmology.  HERA is a compact hexagonal-packed array of 14-meter paraboloids 
-operating from 50 - 225 MHz.
+cosmology.  The current stage under development and described in this paper is a compact hexagonal-packed array of 352 14-meter paraboloids 
+operating from about 70 - 220 MHz.
 \end{abstract}
@@
 \label{sec:intro}
-The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged project
+The Hydrogen Epoch of Reionization Array roadmap (HERA Roadmap) is a staged project
 that uses the unique properties of the 21-cm line of neutral hydrogen to probe the
@@
 Millimeter, and Sub-millimeter category of recommended new facilities in the most
-recent decadal survey of astronomical priorities (\ref{NWNH}). Current projects
-(PAPER, MWA, LOFAR, GMRT YYYYYGIVE REFSYYYYY) are striving to make the first detection of the statistical
+recent Decadal Survey of Astronomy and Astrophysics \citet{NWNH}. Current projects
+(PAPER \citep{2014ApJ...788..106P}, 
+MWA, \citep{2013PASA...30....7T},
+LOFAR \citep{2013A&A...550A.136Y}, 
+GMRT \citep{2011MNRAS.413.1174P}) are striving to make the first detection of the statistical
 power spectrum of the signal, but current best limits still fall above even
@@
 
-The HERA roadmap is a phased set of experiments that build on its predecessors and preceding stages.
+The HERA Roadmap is a phased set of experiments that build on its predecessors and preceding stages.
 The first phase of the HERA roadmap entailed the operation of the
@@
 
-Building on this understanding, the next stage of HERA incorporates a new
+Building on this understanding, the next stage called the Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) incorporates a new
 14m diameter antenna element that is optimized both for sensitivity
@@
 %\appendix
-%\section{\\ \\ \hspace*{-7mm} HEADING}    %% Appendix A
-%And everything nice.
-%\subsection                               %% Appendix A1, A2, etc.
-%And everything else.
+%\section{}    %% Appendix A
+%\subsection{}                               %% Appendix A1, A2, etc.
 
@@
 
+%% REFERENCES
+
 \bibliographystyle{copernicus}
-\bibliography(biblio)
-
-%\begin{thebibliography}{}
-%
-%\bibitem[AUTHOR(YEAR)]{LABEL}
-%REFERENCE 1
-%
-%\bibitem[AUTHOR(YEAR)]{LABEL}
-%REFERENCE 2
-%
-%...
-%
-%\end{thebibliography}
-
-%% Literature citations
+\bibliography{biblio}
+%%
+%% URLs and DOIs can be entered in your BibTeX file as:
+%%
+%% URL = {http://www.xyz.org/~jones/idx_g.htm}
+%% DOI = {10.5194/xyz}
+
+%% LITERATURE CITATIONS
+%%
 %% command                        & example result
-%% \citet{jones90}|               & Jones et al.\ (1990)
+%% \citet{jones90}|               & Jones et al. (1990)
 %% \citep{jones90}|               & (Jones et al., 1990)
@@
 
-%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}[ht]
-\vspace*{2mm}
-\begin{center}
-\includegraphics[width=\textwidth]{plots/herauniall.png}
-\end{center}
+%% FIGURES
+
+%% ONE-COLUMN FIGURES
+
+%f
+\begin{figure}[t]
+\includegraphics[width=8.3cm]{plots/herauniall.png}
 \caption{\small Cartoon showing the development of the Universe and HERA's observing window into it (dashed lines).  The blue solid
@@
 \label{fig:theUniverse}
-\end{figure*}
-
-\begin{figure}[t]\centering
-\includegraphics[width=\columnwidth]{plots/eor_pspec_2014.png}
-\includegraphics[width=\columnwidth]{plots/hera_pk3pk.png}
+\end{figure}
+
+\begin{figure}[t]
+\includegraphics[width=4.1cm]{plots/eor_pspec_2014.png}
+\includegraphics[width=4.1cm]{plots/hera_pk3pk.png}
 \caption{\small Top: HERA's power-spectrum sensitivity (solid)
@@
 constrain the brightness temperature of the IGM at $z\sim8$, showing
-a departure from adiabatic cooling presumed to be indicative of X-ray heating.
-}\label{fig:eor_pspec}
+a departure from adiabatic cooling presumed to be indicative of X-ray heating.}
+\label{fig:eor_pspec}
 \end{figure}
@@
 \begin{figure}[t]
-\vspace*{2mm}
-\begin{center}
-\includegraphics[width=\columnwidth]{plots/heraXY.png}
-\end{center}
-\caption{Scaling  \label{fig:heraXY}}
+\includegraphics[width=8.3cm]{plots/heraXY.png}
+\caption{\small Scaling}
+\label{fig:heraXY}
 \end{figure}
@@
 \begin{figure}[t]
-\vspace*{2mm}
-\begin{center}
-\includegraphics[width=\columnwidth]{plots/costing.png}
-\end{center}
-\caption{Costing  \label{fig:costing}}
+\includegraphics[width=8.3cm]{plots/costing.png}
+\caption{Costing}
+\label{fig:costing}
 \end{figure}
@@
 \begin{figure}[t]
-    \centering
-        \includegraphics[height=1.5in]{plots/new_antenna_closeup.jpg}
-        \includegraphics[height=1.5in]{plots/Engineering/recv_node.png}
-        \includegraphics[height=1.5in]{plots/Engineering/digital.png}
-    \caption{-0.1in}{0.99}{-0.1in}{\small
+        \includegraphics[width=2.7cm]{plots/new_antenna_closeup.jpg}
+        \includegraphics[width=2.7cm]{plots/recv_node.png}
+        \includegraphics[width=2.7cm]{plots/digital.png}
+    \caption{\small
     Existing components re-used in the HERA design include:
@@
     receivers in a node module (center), and
-    the 128-element correlator deployed in the Karoo (right).
-}\label{fig:components}
+    the 128-element correlator deployed in the Karoo (right).}
+\label{fig:components}
 \end{figure}
@@
 
-\begin{figure}[t]
-\vspace*{2mm}
-\begin{center}
+%f
+\begin{figure*}[t]
 \includegraphics[width=12cm]{plots/HERA_high_level_block_diagram.png}
-\end{center}
-\caption{System  \label{fig:systemOverview}}
-\end{figure}
-
-\begin{figure}[t]
-\vspace*{2mm}
-\begin{center}
+\caption{System}
+\label{fig:systemOverview}
+\end{figure*}
+
+\begin{figure*}[t]
 \includegraphics[width=12cm]{plots/config.png}
-\end{center}
-\caption{Configuration  \label{fig:config}}
-\end{figure}
-
-%f
-%\begin{figure*}[t]
-%\vspace*{2mm}
-%\begin{center}
-%\includegraphics[width=12cm]{FILE NAME}
-%\end{center}
-%\caption{TEXT}
-%\end{figure*}
-
-%% TABLES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\caption{Configuration}
+\label{fig:config}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table*}[t]
@@
 
-%% ONE-COLUMN TABLE
-
-%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\vskip4mm
-%\centering
-%\begin{tabular}{column = lcr}
-%\tophline
-
-%\middlehline
-
-%\bottomhline
-%\end{tabular}
-%\end{table}
 \begin{table*}[t]
@@
 
-Print this is some text in between.
-
 \begin{table*}[t]
-\small
-%\begin{center}
-% \begin{deluxetable}{lcc}
-% \tabletypesize{\small}
-% \tablecaption{HERA-331 basic parameters.}
-% \tablehead{\colhead{Parameter} & \colhead{Design} & \colhead{Performance at 150 MHz}}
-%    \startdata
 \begin{tabular}{l | c | c} \hline
@@
 %     ~{   }Point source RMS & 50 $\mu$Jy in 100 hrs \\
-%  \enddata
-%    \hline
- %\end{deluxetable}
 \end{tabular}
@@
 \label{tab:BasicParameters}
-\vspace{.2in}
-%\end{deluxetable}
-%\end{center}
 \end{table*}
 
-%\end{deluxetable}
-
-%% TWO-COLUMN TABLE
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% TABLES
+%%
+%% The different columns must be seperated with a & command and should
+%% end with \\ to identify the column brake.
+
+%% ONE-COLUMN TABLE
 
 %t
-%\begin{table*}[t]
+%\begin{table}[t]
 %\caption{TEXT}
-%\vskip4mm
-%\centering
 %\begin{tabular}{column = lcr}
 %\tophline
-
+%
 %\middlehline
-
+%
 %\bottomhline
 %\end{tabular}
-%\end{table*}
-
-%% The different columns must be seperated with a & command and should
-%% end with \\ to identify the column brake.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%\belowtable{} % Table Footnotes
+%\end{table}
+
+%% NUMBERING OF FIGURES AND TABLES
+%%
 %% If figures and tables must be numbered 1a, 1b, etc. the following command
 %% should be inserted before the begin{} command.
-
 %\addtocounter{figure}{-1}\renewcommand{\thefigure}{\arabic{figure}a}
 
+%% MATHEMATICAL EXPRESSIONS
+
+%% All papers typeset by Copernicus Publications follow the math typesetting regulations
+%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
+%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
+%%
+%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
+%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
+%% Items/objects which are defined are typeset in roman font (Car A, Car B)
+%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
+%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
+%% Vectors are identified in bold italic font using \vec{x}
+%% Matrices are identified in bold roman font
+%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
+%% The character * should not be applied as mutliplication sign
+
 \end{document}

### 2015-08-22 manuscript.tex
update...

@@
 \RequirePackage{ifpdf}
-\documentclass{ICEAA-IEEE_APWC}
+\documentclass{article}
 \usepackage{multirow}
+\usepackage{graphicx}
 
@@
 \author{D. R. DeBoer\thanks{Radio Astronomy Laboratory, University of California, Berkeley, CA 94720, USA, e-mail: ddeboer@berkeley.edu}, 
-on behalf of the HERA collaboration
+team of dozens.
 }
@@
 
-\begin{abstract}
+\section{abstract}
+\label{sec:abstract}
 The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR).
@@
 This paper will provide an overview of the project and describe the design of the HERA receiving element.
-
-\end{abstract}
 
@@
 \begin{figure}[t]
-\centerline{
-\includegraphics[width=7.5cm]{hera331_Bpers.png} }
-\caption{\small Perspective cartoon of the 331-element core within a 305-meter circle.
+\includegraphics[width=7.5cm]{hera331_Bpers.png}
+\caption{Perspective cartoon of the 331-element core within a 305-meter circle.
 \label{fig:config}}
@@
 
-HERA is a collaboration of the following partner institutions:  Arizona State University, University of California Berkeley, University of California Los Angeles, University of Cambridge, University of KwaZulu Natal, Massachusetts Institute of Technology, National Radio Astronomy Observatory, University of Pennsylvania, SKA-South Africa, and
+HERA is a collaboration of the following partner institutions:  Arizona State University, Brown University, University of California Berkeley, University of California Los Angeles, University of Cambridge, Massachusetts Institute of Technology, National Radio Astronomy Observatory, University of Pennsylvania, SKA-South Africa, and
 University of Washington.
@@
 The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
-Frequency therefore equates to a volume along the site of the telescope as well as in dating when the event occurred.
+Frequency therefore equates to a cosmic volume along the site of the telescope, as well as in dating when the event occurred.
 
@@
 Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of its properties to achieve this more difficult goal.
+There are now numerous papers in the literature the technique of measuring the power spectrum using an interferometer and the attendant issues (list of citations).
+Below, a high-level overview based on [parsons12b] will be presented.
 
@@
 \begin{equation}
-|\kpar| \le \frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
+|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
 \end{equation}
@@
 
-\section{HERA System}
+\section{Specifications}
+\label{sec:specs}
+Based on the discussion above, we see that
+
+\section{HERA System Overview}
 \label{sec:system}
@@
 
-\section{HERA Element}
+\subsection{HERA Element}
 \label{sec:element}

### 2015-11-03 manuscript.tex
barebones version

@@
 % refer to the same affiliation of the n:th \thanks command
-\author{D. R. DeBoer\thanks{Radio Astronomy Laboratory, University of California, Berkeley, CA 94720, USA, e-mail: ddeboer@berkeley.edu}, 
-team of dozens.
+\author{A.U. Thor\thanks{Institution}
 }
@@
 
-\section{abstract}
+\section{Abstract}
 \label{sec:abstract}
@@
 In our analogy, this is akin to following our child's activities and growth through a good part of their formative years.
+Fig \ref{fig:cosmos}.
+
+\begin{figure}[t]
+\includegraphics[width=\textwidth]{plots/herauni/herauniall.png}
+\caption{History of the Universe  CREDIT BACKGROUND IMAGE - GET PERMISSION
+\label{fig:cosmos}}
+\end{figure}
 
@@
 \begin{figure}[t]
-\includegraphics[width=7.5cm]{hera331_Bpers.png}
-\caption{Perspective cartoon of the 331-element core within a 305-meter circle.
-\label{fig:config}}
+\includegraphics[width=0.45\textwidth]{plots/HERA_array.png}
+\includegraphics[width=0.45\textwidth]{plots/IMG_3538.png}
+\caption{Perspective cartoon of the 331-element core within a 305-meter circle.  With current picture.
+\label{fig:array}}
 \end{figure}
@@
 This has resulted in a close-packed array of transit 14-meter dishes (Fig.
-\ref{fig:config}), with the first elements currently under construction at the South African Karoo Astronomy Reserve, the current location of the Donald C.
+\ref{fig:array}), with the first elements currently under construction at the South African Karoo Astronomy Reserve, the current location of the Donald C.
 Backer Precision Array to Probe the Epoch of Reionization (PAPER) array.
@@
 The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
-Frequency therefore equates to a cosmic volume along the site of the telescope, as well as in dating when the event occurred.
+The bandwidth
+therefore equates to a cosmic volume along the site of the telescope, with the frequency determining the cosmic age.
 
@@
 The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
-Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of its properties to achieve this more difficult goal.
+Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.
 There are now numerous papers in the literature the technique of measuring the power spectrum using an interferometer and the attendant issues (list of citations).
@@
 
-Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment (see {\em e.g.} \cite{2012ApJ...756..165P}):
-\begin{equation}
-{\hat P}(\kvec) \approx \left(\frac{X^2Y}{4k_B^2}\right) \frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} 
-\end{equation}
-where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated squared beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
-The terms not in parentheses are instrumental terms, as opposed to constants and cosmological parameters.
-
-Though the desired measurement is over a 3-D volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
+Though the desired measurement is over a volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
 This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything not the EOR) contaminate the signal of interest from where they don't.
-This contaminated phase space may be shown to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
 \begin{equation}
@@
 \end{equation}
-where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response.  
+where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
 A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
@@
 
+\begin{figure}[t]
+\includegraphics[width=\textwidth]{plots/wedge.png}
+\caption{Cartoon of the wedge.
+The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
+The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
+The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
+The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.
+\label{fig:wedge}}
+\end{figure}
+
+The techniques to measure the power spectrum may be broken down into two principle techniques, plus additional hydrid methods.  These are briefly summarized below.
+
+\subsection{Delay-Spectrum Approach}
+\label{sec:delayapproach}
+Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
+The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
+Other than handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
+
+We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment (see {\em e.g.} \cite{2012ApJ...756..165P}):
+\begin{equation}
+{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
+\label{eq:Pk_baseline}
+\end{equation}
+where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
+The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
+Rather than ${\hat P}(\kvec)$, the literature generally works with a dimensionless parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+
+Since the thermal noise per visibility baseline may be expressed as
+\begin{equation}
+V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
+\label{eq:sensitivity_per_baseline}
+\end{equation}
+where $\tau$ is the integration time (see {\em e.g.} TMS),
+to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
+\ref{eq:Pk_baseline}.
+Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
+\begin{equation}
+\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
+\end{equation}
+where $\mathcal{N}$ represents the improvement in sensitivity for coherently and incoherently averaging $k$-bins, depending on the configuration and technique (see \ref{parsons12b}).
+
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
+Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
+Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
+The red-dashed line is the allowed smallest k-value that can be measured.
+Note that at these red-shifts it is essentially set just by the bandwidth.
+
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=\textwidth]{plots/kperf_heraZ.png} 
+}
+\caption{\small k
+\label{fig:kperf}}
+\end{figure}
+
+\subsection{Map-Making Approach}
+\label{sec:mapapproach}
+In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
+The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum.
+Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller k-modes on the sky.
+The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
+In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
+
+\subsection{Hybrid Approaches}
+\label{sec:hybridapproach}
+
 \section{Specifications}
 \label{sec:specs}
-Based on the discussion above, we see that
+As the initial detection and analysis approach is primarily focused on the delay-spectrum approach, the requirements are set primarily by that approach.
+However, since the map-making approaches remain a goal, the design makes every effort to allow such studies.
+The high-level specifications are summarized in Fig.
+\ref{fig:array}.
+Specifications fall within one of three categories:  Phase 1, Phase 2, or Extended.
+
+\begin{enumerate}
+\item {\em Phase 1} corresponds to the current stage where the existing analog and digital signal chains are used with the new HERA element.
+Ostensibly, this corresponds to the first two years where the total number of HERA elements remains less than 128.
+This allows for a strong transition period with a known signal path but with new elements.
+\item {\em Phase 2} corresponds to the upgraded signal path slated for when the element count exceeds 128 and the architecture changes to allow extensibility and improved performance.
+\item {\em Extended} are those goals that are aspirational on a best efforts basis, or that include additional science beyond the EOR.
+\end{enumerate}
+
+\subsection{Scientific Requirements and Specifications}
+\subsubsection{Spectral Response}
+Figure \ref{fig:cosmos} highlights the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
+These limits are derived from to-date complementary probes of reionization which include measurements of
+the optical depth to last scattering in the CMB, QSO spectra, Ly-$\alpha$
+absorption in the spectra of quasars and gamma ray bursts and the demographics
+of Ly-$\alpha$ emitting galaxies
+%We show limits established by some of these techniques in
+(Figure~\ref{fig:IonHist}). 
+Constraints from these probes are still weak:
+Ly-$\alpha$ absorption saturates at very small neutral fractions; galaxy
+surveys directly constrain only the bright end of the luminosity function and
+depend on an unknown escape fraction of ionizing photons to constrain
+reionization; CMB measurements probe an integral quantity subject
+to large degeneracies. Even when these observations are
+combined into a single 95\% confidence region, the bounds remain weak.
+For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
+at $z=8$. 21\,cm reionization experiments place much tighter constraints on ionization, with the
+red band showing the forecasted 95\% confidence region derived from HERA data,
+after marginalizing over astrophysical and cosmological parameters.
+
+****Include references
+
+From these measurements and models, the Phase 1 and 2 minimum spans of redshifts range from $z=$6 to 13, corresponding to a frequency range of 100 - 200 MHz.
+Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to about 50 MHz without compromising the performance in the specification range.
+The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.
+The 100 MHz bandwidth is also the minimum fully digitized and correlated bandwidth.
+
+The scientific requirement on channel bandwidth is to allow access to k-modes of XXX 5?
+h/Mpc [ref], which is easily accomplished by as few as 64 channels over the 100 MHz bandwidth.
+However, in order to handle radio frequency interference as well as to allow the bandwidth to be characterized a specification of 1024 channels has been set.
+This yields a channel bandwidth of 97.7 kHz.
+
+\subsubsection{Spatial/Delay Response}
+The spatial response provides the defining characteristics for the experiment; related to the diameter, focal length, baseline, and configuration.
+The principal constraint sets the delay response of the antenna element, characterized by the magnitude and frequency of the antenna ripple response.
+This was initially addressed to first order in [MEMO] and in a much more detailed way in [UPCOMING PAPERS].
+Expressed simply, the delay response of the antenna must be sufficiently attenuated such that the foreground signal levels at those delays is below the expected signal strength of the EOR signal.
+This sets the optics and match needs.
+As discussed in [REFS], this is addressed by direct measurement and modeling of the response to the system relative to and as a function of delay time from the initial wave reception.
+The expected strength of the foregrounds is modeled through the system and plotted on the same set of axes.
+These two regions should be disjoint to a significance of 60 dB.
+
+Spatial modes - delay 60@60, baseline, diameter
+
+\subsubsection{Sensitivity}
+sensitivity - reionization fraction 10$\sigma$ from 0.2-0.8
+
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=.3\textwidth]{plots/ionHist.pdf}
+\includegraphics[width=.3\textwidth]{plots/eor_pspec_2014.png}
+\includegraphics[width=.3\textwidth]{plots/HERA_sensitivity.pdf}
+}
+\caption{\small Sensitivity
+\label{fig:sensivity}}
+\end{figure}
+
+\subsection{Programmatic Specifications}
+stages/soon - phase with initial PAPER equipment
+
+cost - cheap experiment
 
@@
 \label{sec:system}
+STAGED APPROACH WITH PAPER FIRST
+
 HERA will operate between 70 MHz and 220 MHz, however the first phases (up to 127 elements) will operate within the current PAPER bandwidth of 110-190 MHz.
@@
 \centerline{
-\includegraphics[width=5.5cm]{system.png} }
+\includegraphics[width=\textwidth]{plots/sysOver.png} }
+\caption{\small Architecture of HERA.
+\label{fig:architecture}}
+\end{figure}
+
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=\textwidth]{plots/HERA_high_level_block_diagram.png} }
 \caption{\small High-level block diagram of the HERA system.
@@
 
-\subsection{HERA Element}
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=7.5cm]{plots/costfigNew.png} 
+}
+\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
+The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
+\label{fig:costfig}}
+\end{figure}
+
+\subsection{Element and Analog Signal Path}
 \label{sec:element}
@@
 \begin{equation}
-N_a = 26918D_a^{-\frac{5}{3}}
+N_a = 4634/D_a
 \end{equation}
@@
 
-\begin{figure}[t]
-\centerline{
-\includegraphics[width=7.5cm]{costfig.png} 
-}
-\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
-The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
-\label{fig:costfig}}
-\end{figure}
-
 A full-sized prototype of this system was built and measured using a network analyzer in time-domain mode, which meets the ``60 by 60 '' specification (Fig \ref{fig:refl}). (\cite{heraMemo5}).
@@
 \centerline{
-\includegraphics[width=7.5cm]{reflections.png} 
+\includegraphics[width=\textwidth]{plots/compareSimToDataNormalized.pdf} 
 }
-\caption{\small Network analyzer measurements of reflected power versus delay time of the full-size HERA prototype.
+\caption{\small Simulations/measurements of reflected power versus delay time of the full-size HERA prototype.
 \label{fig:refl}}
 \end{figure}
+
+\subsection{Digital Signal Path}
+\label{sec:DSP}
+
+\subsection{Real-Time Processor and Librarian}
+\label{sec:rtplib}
+
+\subsection{Monitor and Control}
+\label{sec:moncon}
+
+\subsection{Infrastructure}
+\label{sec:infra}
+
+\subsection{Post-Processing}
+\label{sec:postproc}
+
+\subsection{Calibration}
 
@@
 \centerline{
-\includegraphics[width=7.5cm]{eorsens.png} 
+\includegraphics[width=7.5cm]{plots/eorsens.png} 
 }

### 2015-11-03 eorsens.py

@@ -1,49 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-sens = """PAPER 128  &  1.56  &  -  &  4.46  &  - 
-MWA 128      &  0.66  & 0.86  &  2.50  &  3.15 
-LOFAR (core) & 0.70  & 1.90  & 7.48  &  12.22 
-HERA 37         & 5.67  &  -  & 15.46   &  - 
-HERA 331       & 38.75 & -  &  111.69  &  - 
-MWA 256         &2.40   &  2.81  &  8.28  & 9.64 
-SKA1 Low      &21.23   & 26.92  & 139.07  & 115.13"""
-
-sens = sens.split('\n')
-arrnames = []
-arrsens = []
-for line in sens:
-    data=line.split('&')
-    arrnames.append(data[0].strip())
-    del(data[0])
-    varr = []
-    for v in data:
-        if v.strip()=='-':
-            varr.append(0.0)
-        else:
-            varr.append(float(v))
-    arrsens.append(varr)
-arrsens = np.array(arrsens)
-print arrnames
-print arrsens
-
-gridlines = [0.6,0.7,0.8,0.9,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,20.0,30.0,40.0,50.0,60.0,70.0,80.0,90.0,100.0]
-gridlines = [1.0,10.0,100.0]
-for g in gridlines:
-    plt.plot([-1.5,13.5],[g,g],'k:')
-
-w = 0.4
-c = ['b','c','m','r']
-labels = ['Avoid(drift)','Subtract(drift)','Avoid(track)','Subtract(track)']
-cols = [0,2,1,3]
-ll = np.arange(len(arrsens[:,0]))*2-0.5+w/2.0
-for i in range(len(cols)):
-    plt.bar(ll+i*w*1.05,arrsens[:,cols[i]],color=c[i],width=w,label=labels[i])
-plt.yscale('log')
-plt.axis([-1.5,13.5,0.5,140.0])
-plt.legend(loc='upper left')
-xt = [0,2,4,6,8,10,12]
-plt.xticks(xt,arrnames,rotation=70)
-plt.plot([5.55,5.55],[0.5,140],'k',linewidth=3)
-
-

### 2015-11-04 manuscript.tex
update to skeleton

@@
 Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.
-This paper will provide an overview of the project and describe the design of the HERA receiving element.
+This paper will provide an overview of the project and describe the design of the system.
 
@@
 In our analogy, this is akin to following our child's activities and growth through a good part of their formative years.
-Fig \ref{fig:cosmos}.
+Fig.
+\ref{fig:cosmos} shows a depiction of the history of the Universe, as well as the 21-cm Hydrogen line red-shifted to cosmic time.
 
@@
 \includegraphics[width=\textwidth]{plots/herauni/herauniall.png}
-\caption{History of the Universe  CREDIT BACKGROUND IMAGE - GET PERMISSION
+\caption{history of the Universe plotted on a log scale.
+The top axis shows the time elapsed since the Big Bang while the bottom shows the corresponding redshift.
+Of course, placing exactly when the depicted events occurred and how is the goal of HERA.
+The right axis and the red-line show the frequency of the 121-cm Hydrogen line over cosmic time.
+The horizontal lines denote the frequency range of HERA, both the initial band (the inner pair) and the desired extended band.
+CREDIT BACKGROUND IMAGE - GET PERMISSION
 \label{fig:cosmos}}
@@
 HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection.
-The key is to produce inexpensive collecting area optimized for the spatial scales needed to detect the EOR.
+The key is to produce inexpensive collecting area optimized to detect the EOR.
 This has resulted in a close-packed array of transit 14-meter dishes (Fig.
@@
 Backer Precision Array to Probe the Epoch of Reionization (PAPER) array.
+[PAPER REF]
 
-Beginning with the current NSF funding, HERA will rollout in a staged deployment.
+Beginning with the current NSF funding, HERA is rolling out in a staged deployment.
 The first phase extending to August 2016 will deploy 37 elements that will test the HERA element design for performance and manufacturability at location.
@@
 
-Initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.
+Given the difficulty of the experiment, initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.
 The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
 Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.
-There are now numerous papers in the literature the technique of measuring the power spectrum using an interferometer and the attendant issues (list of citations).
-Below, a high-level overview based on [parsons12b] will be presented.
+
+Obviously between our present observation point and the Epoch of Reionization approximately 13 billion years ago lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
+As a first step, areas of the sky with where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
+More importantly and most fortuitously however is the fact that all of these signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
+This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.
 
@@
 Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
+There are now numerous papers in the literature the technique of measuring the power spectrum using an interferometer and the attendant issues (list of citations).
+Below, a high-level overview based on  \cite{2012ApJ...756..165P} will be presented.
 The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
@@
 
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment (see {\em e.g.} \cite{2012ApJ...756..165P}):
+We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
 \begin{equation}
@@
 The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Rather than ${\hat P}(\kvec)$, the literature generally works with a dimensionless parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
 
@@
 
-\section{Specifications}
+\section{Scientific Requirements and Specifications}
 \label{sec:specs}
@@
 
-\subsection{Scientific Requirements and Specifications}
-\subsubsection{Spectral Response}
+\subsection{Spectral Response}
 Figure \ref{fig:cosmos} highlights the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
@@
 
-\subsubsection{Spatial/Delay Response}
+\subsection{Spatial/Delay Response}
 The spatial response provides the defining characteristics for the experiment; related to the diameter, focal length, baseline, and configuration.
@@
 
-Spatial modes - delay 60@60, baseline, diameter
+The minimum baseline is set such that the baseline has negligible impact on the minimum baseline (shown in Fig.
+\ref{fig:kperf} in the standard analysis.
+Somewhat arbitrarily, a maximum shortest baseline with equal contribution to $k_{min}$ at the full bandwidth of 100 MHz is chosen, which corresponds to about 15m.
+Additionally, this baseline has been well-studied by the PAPER array (ref).
 
-\subsubsection{Sensitivity}
+The above requirements also drive the specification on the element diameter, as well as the desire to maximize sensitivity.
+This means that the diameter should be something around 15m (that is, a fully filled aperture for the minimum baseline) with a short focal length.
+
+\subsection{Sensitivity}
 sensitivity - reionization fraction 10$\sigma$ from 0.2-0.8
@@
 
-\subsection{Programmatic Specifications}
-stages/soon - phase with initial PAPER equipment
-
-cost - cheap experiment
+\subsection{Cost}
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=7.5cm]{plots/costfigNew.png} 
+}
+\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
+The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
+\label{fig:costfig}}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-\centerline{
-\includegraphics[width=7.5cm]{plots/costfigNew.png} 
-}
-\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
-The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
-\label{fig:costfig}}
-\end{figure}
-
 \subsection{Element and Analog Signal Path}
@@
 
+\subsection{Configuration}
+\label{sec:config}
+
 \subsection{Calibration}

### 2015-11-04 sens_discuss.tex
update to skeleton

\documentclass[11pt]{article}

\newcommand{\kvec}{{\bf k}}
\newcommand{\bvec}{{\bf b}}
\newcommand{\shat}{{\hat s}}
\newcommand{\kpr}{{k_\perp}}
\newcommand{\kvpr}{{\kvec_\perp}}
\newcommand{\kpl}{{k_\parallel}}
\newcommand{\AI}{{\langle\tilde A*\tilde I\rangle}}
\newcommand{\AItau}{{\AI(\tau)}}
\newcommand{\hMpci}{{h~{\rm Mpc}^{-1}}}
\newcommand{\inch}{$^{\prime\prime}$}
\newcommand{\foot}{$^{\prime}$}
\renewcommand{\deg}{^\circ}
\begin{document}

From Parsons {\em et al} 2012:

\small
\begin{equation}
\label{eq:sensitivity}
\Delta^2_N(k) \propto (k)^\frac{5}{2}
                                         \left[\frac{1}{B}\right]^\frac{1}{2}
                                         \left[\frac{1}{\Delta\ln k}\right]^\frac{1}{2} \\
                           \Omega T_{sys}^2
                                         \left[\frac{1}{t_{per\_day}}\right]^\frac{1}{2}
                           \left[\frac{1}{t_{cam}}\right]
                                         \left[\frac{1}{N_a}\right]
                                         \left[\frac{f_o}{f}\right]^\frac{1}{2}
\end{equation}
where $k$ is the magnitude of the $k$-mode, $B$ is the bandwidth, $\Delta\ln k$ is the log
of the binsize, $\Omega$ is the field-of-view, $T_{sys}$ is the system temperature, 
${t_per\_day}$ is the number of hours observed per day, $t_{cam}$ is the number of days
observed, $N_a$ is the number of antennas, and $f_o/f$ is a configuration metric for a 
redundant array as therein defined.

Pulling out terms relating to diameter ($D_a$) and number, we can write Eq. \ref{eq:sensitivity} as
\begin{equation}
\label{eq:reducedSensitivity}
\Delta^2_N(k) \propto \frac{\Omega \sqrt{f_o/f}}{N_a\sqrt{t_{per\_day}}} \propto \frac{(1/D_a^2)(1/\sqrt{N_a})}{N_a\sqrt{D_a}}
= \frac{1}{D_a^{\frac{5}{2}}N_a^{\frac{3}{2}}}
\end{equation}
where the dependencies on diameter and number have been substituted in, noting that the expressions for $f_o/f$ and 
$t_per\_day$ were derived in Parsons et al where the baselines for the close-packed array are multiples of the diameter.

From Mellema {\em et al} 2013:

\begin{equation}
\Delta^2_N(k) \propto k^\frac{3}{2}\sqrt{D_c^2\Delta D_c\Omega_{FOV}}\left[\frac{T_{sys}}{\sqrt{Bt_{int}}}\right]^2\left[\frac{A_{core}A_{eff}}{A^2_{coll}}\right]
\end{equation}

Pulling out similar $D_a$ and number terms:

\begin{equation}
\Delta^2_N(k) \propto \sqrt{\Omega} \left[ \frac{1}{t_{int}}  \right] \left[\frac{A_{core}A_{eff}}{A^2_{coll}}\right] \propto
\left[\frac{1}{D_a}\right] \left[ \frac{1}{t_{int}}  \right] \left[\frac{(ND_a^2)(D_a^2)}{(ND_a^2)^2}\right] = \frac{1}{t_{int}}\frac{1}{ND_a^2}
\end{equation}

where, for hex packing
\begin{equation}
A_{core} \propto \left[\frac{4N-1}{3}D_a^2\right] \sim ND_a^2
\end{equation}

So, there is a factor of $\sqrt{ND_a}$ different.
Allowing for incoherent drift-scanning, $t_{int} \rightarrow \sqrt{t_{int}} \sim 1/ \sqrt{D_a}$, so we are left with a $\sqrt{N}$ different, probably related to the specifics of $f_o/f$ and/or $A_{core}/A_{coll}^2$...

Running Jonnie's code for 
\begin{itemize}
\item a bunch of diameters with 331 antennas (7-21 m)
\item a bunch of elements for 14-m diameter (37-919)
\end{itemize}
 and evaluate at ks = 1, yields:
\begin{equation}
\Delta^2_N(1) \approx \frac{1}{D^{1.5}N^{1.5}}
\end{equation}
The exact coefficient is weakly dependent on the k-value in the middle k values.  The power laws fit better for larger k's (and are closer to 1.5).

\end{document}

### 2015-11-04 plots/wedge.py

@@ -1,0 +1,44 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+kperp = np.logspace(-3,-1,100)
+kperp = np.linspace(0,.1,100)
+kzero = np.zeros(len(kperp))
+kflat = np.ones(len(kperp))*0.1
+plt.plot(kperp,kflat)
+plt.fill_between(kperp,kzero,kflat,color='k',alpha='0.5')
+
+khoriz = 5.0*kperp
+g = np.where(khoriz < max(kflat))
+khoriz[g] = max(kflat)
+plt.plot(kperp,khoriz)
+plt.fill_between(kperp,kflat,khoriz,color='b',alpha='0.5')
+#1ksys = khoriz + 0.05 + 1.1e-3/(kperp+1e-4)
+#1ksys[g] = ksys[g[0][-1]+1]*np.ones(len(g[0]))
+#2ksys = 4.1*(kperp+0.00) + 0.15
+x = [0.0,0.01,0.09]
+y = [0.15, 0.18, 0.50]
+num = (y[2]-y[0]) - (y[1]-y[0])*(x[2]**2-x[0]**2)/(x[1]**2-x[0]**2)
+den = (x[2]-x[0]) - (x[1]-x[0])*(x[2]**2-x[0]**2)/(x[1]**2-x[0]**2)
+B = num/den
+A = (y[1]-y[0] - B*(x[1]-x[0]))/(x[1]**2-x[0]**2)
+C = y[0] - A*x[0]**2 - B*x[0]
+ksys = A*kperp**2 + B*kperp + C
+plt.plot(kperp,ksys,'r--')
+plt.fill_between(kperp,khoriz,ksys,color='r',alpha='0.2')
+plt.axis([0.0,0.1,.0,.55])
+plt.xlabel(r'k$_\perp$',fontsize=14)
+plt.ylabel(r'k$_{||}$',fontsize=14)
+
+#cables
+kperp_cables = [0,0.085]
+kpam = 0.45
+kpar_cables1 = [kpam,kpam]
+plt.plot(kperp_cables,kpar_cables1,'r--')
+kpam = 0.465
+kpar_cables2 = [kpam,kpam]
+plt.plot(kperp_cables,kpar_cables2,'r--')
+plt.fill_between(kperp_cables,kpar_cables1,kpar_cables2,color='r',alpha='0.2')
+
+
+

### 2015-11-04 manuscript.tex
fixed error in channel number, updated wedge fig and added config fig

@@
 \begin{figure}[t]
-\includegraphics[width=\textwidth]{plots/wedge.png}
+\includegraphics[height=3in]{plots/wedge.png}
 \caption{Cartoon of the wedge.
@@
 The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
 
@@
 
-The scientific requirement on channel bandwidth is to allow access to k-modes of XXX 5?
-h/Mpc [ref], which is easily accomplished by as few as 64 channels over the 100 MHz bandwidth.
-However, in order to handle radio frequency interference as well as to allow the bandwidth to be characterized a specification of 1024 channels has been set.
+The scientific requirement on channel bandwidth is to allow access to k-modes of greater than 1 h/Mpc [ref], which requires about 256 channels over the 100 MHz bandwidth.
+However, in order to handle radio frequency interference as well as to allow the bandpass to be characterized, a specification of 1024 channels has been chosen.
 This yields a channel bandwidth of 97.7 kHz.
@@
 
-The minimum baseline is set such that the baseline has negligible impact on the minimum baseline (shown in Fig.
-\ref{fig:kperf} in the standard analysis.
+The minimum baseline is set such that the baseline has negligible impact on the minimum $k_{bin}$ (shown in Fig.
+\ref{fig:kperf}) in the standard analysis.
 Somewhat arbitrarily, a maximum shortest baseline with equal contribution to $k_{min}$ at the full bandwidth of 100 MHz is chosen, which corresponds to about 15m.
@@
 This means that the diameter should be something around 15m (that is, a fully filled aperture for the minimum baseline) with a short focal length.
+This is revisited in Sec \ref{sec:cost}.
 
@@
 \subsection{Cost}
+\label{sec:cost}
 \begin{figure}[t]
@@
 \label{sec:config}
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=7.5cm]{plots/config.png} 
+}
+\caption{\small Config
+\label{fig:config}}
+\end{figure}
 

### 2015-11-04 manuscript.tex
many edits in the spatial response section

@@
 
-\subsection{Spectral Response}
+\begin{table}
+\label{tab:specs}
+\caption{Overview of requirements and specifications}
+\begin{tabular}{| l | l | c |}\hline
+Parameter & Value & Comments \\ \hline
+\end{tabular}
+\end{table}
+
+\subsection{Spectral Response:  Frequency Range and Bandwidths}
 Figure \ref{fig:cosmos} highlights the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
@@
 
-\subsection{Spatial/Delay Response}
+\subsection{Spatial/Delay Response:  Diameter and Baselines}
 The spatial response provides the defining characteristics for the experiment; related to the diameter, focal length, baseline, and configuration.
@@
 
+In order to do an initial design and a parametric study, the approach used first sets a specification on the magnitude of the feed/vertex reflections within the dish for 1, 2, and 3 round-trip travel delay times.
+Reference \cite{heraMemo5} provides an initial estimate that the reflections must be down by 60 dB on timescales greater then 60 nanoseconds.
+Also in ref \cite{heraMemo5}, an analysis using a feed pattern based on the existing PAPER dipole antenna with a fairly complete analytical model of a prime focus antenna, the effective area was found to be maximized for $f/D \approx 0.32$.
+Figure \ref{fig:costfig} shows the vertex/feed round-trip travel times for one, two and three reflections.
+
+Interpreting these delay lines in Fig \ref{fig:costfig} is somewhat confusing, but note that the vertical color scale represents a fairly steep cost function preferring delay time-scales of less than 60 ns (ignoring the horizontal color scale, to be discussed in Sec \ref{sec:cost}).
+The preliminary study requirement is to ensure that reflections are down by about 60 dB by that time-scale.
+As seen by the black dashed line labelled '1' for one round-trip between the feed and vertex, for the focal ratio and diameters of interest  this reflection is always less than 60 ns, so this does not constrain the design.
+For two round-trips (black solid line labelled '2'), we see that for diameters greater than 14.1m, the delay is less than 60 ns, so diameters less than this are preferred.
+For three round-trips, we see that we would need antennas with diameters smaller than 9.4m.
+However, cost (sec.
+\ref{sec:cost) and practicality issues ({\em i.e.}physically moving within the filled array) make these small diameters undesirable, so this means that we need to ensure that our system has sufficiently damped reflections for 3 reflections.
+This is discussed in detail in UPCOMING PAPERS.
+
 The minimum baseline is set such that the baseline has negligible impact on the minimum $k_{bin}$ (shown in Fig.
@@
 The above requirements also drive the specification on the element diameter, as well as the desire to maximize sensitivity.
-This means that the diameter should be something around 15m (that is, a fully filled aperture for the minimum baseline) with a short focal length.
-This is revisited in Sec \ref{sec:cost}.
+This means that the diameter should be something around 15m (that is, a fully filled aperture for the minimum baseline).
 
-\subsection{Sensitivity}
-sensitivity - reionization fraction 10$\sigma$ from 0.2-0.8
+\subsection{Sensitivity:  Diameter, Number, and Configuration}
+What is our requirement?
+
+sensitivity - reionization fraction 10$\sigma$ from 0.2-0.8?
+
+10\% estimate of neutral fraction?
 
@@
 The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
+SEE TEXT...
 \label{fig:costfig}}
@@
 And the third items constrains the construction materials and methods and the operational model.
-
-The approach used first sets a specification on the magnitude of the reflections within the dish on the delays of interest.
-Reference \cite{heraMemo5} derives this to be that the reflections must be down by 60 dB on timescales greater then 60 nanoseconds.
-Next, using a feed pattern based on the existing PAPER dipole antenna and a fairly complete analytical model of a prime focus antenna, the effective area was found to be maximized for $f/D \approx 0.32$.
-Figure \ref{fig:costfig} shows the vertex/feed round-trip travel times for one, two and three reflections with the 60 ns timescale indicated by the black dashed line.
 

### 2015-11-04 manuscript.tex
incremental but added specs table

@@
 \caption{Overview of requirements and specifications}
-\begin{tabular}{| l | l | c |}\hline
-Parameter & Value & Comments \\ \hline
+\begin{tabular}{| l | l | l | p{5cm} |}\hline
+{\bf Parameter} & {\bf Requirement} & {\bf Value} & {\bf Comments} \\ \hline
+Frequency range & 100 - 200 MHz & & Phase 1 \\ \hline
+Number of channels & $>$256 & 1024 & \\ \hline
+Diameter    & $<$ 14.1 m &  14.0 m & Picked near requirement for sensitivity/cost \\ \hline
+Shortest baseline & $<$ 15 m & 14.6 m & \\ \hline
 \end{tabular}
@@
 
-\subsection{Spatial/Delay Response:  Diameter and Baselines}
+\subsection{Spatial/Delay Response:  Diameter and Minimum Baseline}
 The spatial response provides the defining characteristics for the experiment; related to the diameter, focal length, baseline, and configuration.
@@
 
-Interpreting these delay lines in Fig \ref{fig:costfig} is somewhat confusing, but note that the vertical color scale represents a fairly steep cost function preferring delay time-scales of less than 60 ns (ignoring the horizontal color scale, to be discussed in Sec \ref{sec:cost}).
+Interpreting Fig \ref{fig:costfig} is somewhat confusing, but note that the vertical color scale represents a fairly steep cost function preferring delay time-scales of less than 60 ns (ignoring the horizontal color scale for now, to be discussed in Sec \ref{sec:cost}).
 The preliminary study requirement is to ensure that reflections are down by about 60 dB by that time-scale.
@@
 However, cost (sec.
-\ref{sec:cost) and practicality issues ({\em i.e.}physically moving within the filled array) make these small diameters undesirable, so this means that we need to ensure that our system has sufficiently damped reflections for 3 reflections.
+\ref{sec:cost} and practicality issues ({\em i.e.}physically moving within the filled array) make these small diameters undesirable, so this means that we need to ensure that our system has sufficiently damped reflections for 3 reflections.
 This is discussed in detail in UPCOMING PAPERS.
+We have therefore adopted a specification that the antenna must be less than 14.1 meters in diameter.
 
@@
 Somewhat arbitrarily, a maximum shortest baseline with equal contribution to $k_{min}$ at the full bandwidth of 100 MHz is chosen, which corresponds to about 15m.
-Additionally, this baseline has been well-studied by the PAPER array (ref).
+Additionally, this baseline has been well-studied by the PAPER array (ref) and is consistent with the diameter specification derived above.
 
-The above requirements also drive the specification on the element diameter, as well as the desire to maximize sensitivity.
-This means that the diameter should be something around 15m (that is, a fully filled aperture for the minimum baseline).
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=7.5cm]{plots/costfigNew.png} 
+}
+\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
+The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
+SEE TEXT...
+\label{fig:costfig}}
+\end{figure}
 
@@
 \label{sec:cost}
-\begin{figure}[t]
-\centerline{
-\includegraphics[width=7.5cm]{plots/costfigNew.png} 
-}
-\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
-The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
-SEE TEXT...
-\label{fig:costfig}}
-\end{figure}
 

### 2015-11-04 manuscript.tex
merged sensitivty and costs spec sections

@@
 
-\subsection{Sensitivity:  Diameter, Number, and Configuration}
+\subsection{Sensitivity and Cost:  Diameter, Number, and Configuration}
+\label{sec:cost}
 What is our requirement?
@@
 \caption{\small Sensitivity
-\label{fig:sensivity}}
+\label{fig:sensitivity}}
 \end{figure}
 
-\subsection{Cost}
-\label{sec:cost}
+To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity for this measurement, which is shown in Figure \ref{fig:costfig}.
+This included the full system, except the post-processing.
+By assuming canonical values of 331 14-meter antennas, the number of elements needed for a given diameter to match the sensitivity is
+\begin{equation}
+N_a = 4634/D_a
+\end{equation}
+where $N_a$ and $D_a$ are the number of antennas and their diameter respectively.
+Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.
+This dependency differs from Mellema, mostly likely due to the assumed configuration.
+Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
+which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
 
@@
 \label{sec:element}
-The design of the HERA element is discussed in \cite{heraMemo5}.  The design principles are three-fold:
+As discussed in the previous section, the design of the design principles are three-fold:
 \begin{itemize}
@@
 And the third items constrains the construction materials and methods and the operational model.
-
-To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity for this measurement, which is shown in Figure \ref{fig:costfig}.
-By assuming canonical values of 331 14-meter antennas, the number of elements needed for a given diameter to match the sensitivity is
-\begin{equation}
-N_a = 4634/D_a
-\end{equation}
-Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve, which becomes quite flat for diameters greater than about 14m, which also happens to be the diameter where two reflections between the feed and vertex coincide with the 60 ns specification.
-An addition high-level analysis of reflection coefficients within the structure also indicate that a diameter of less than 15 meters is preferred.
-The diameter of 14 meters was therefore chosen for the element.
-
-A full-sized prototype of this system was built and measured using a network analyzer in time-domain mode, which meets the ``60 by 60 '' specification (Fig \ref{fig:refl}). (\cite{heraMemo5}).
+This led to an element diameter of 14 meters.
 
